---
title: "NMDS Analysis of Microbial Communities"
output: html_document
---

# Load Libraries
```{r}
library(tidyverse)
library(vegan)
library(ggplot2)
```

# Step 1: Read the data
```{r}
soil_chem_data <- read.csv("/Users/kyasparks/Desktop/Desktop - MacBook Pro (2)/Kyas_PhD/CSU_2023_2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/Soil_Chemistry_TimFegelRMRS/SoilChemAnalysis_16S_ASCC/ASCC_16S_SoilChemMetadata_rem.csv")


feature <- read.delim("/Users/kyasparks/Desktop/Desktop - MacBook Pro (2)/Kyas_PhD/CSU_2023_2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_metadata/ASCC_16S_feature_table_rem_RELABUND.txt")
```


# Step 2: Reshape the feature data into a long format
```{r}
microbe_data_long <- feature %>%
  pivot_longer(cols = 4:306, names_to = "Sample", values_to = "Abundance") %>%
  replace(is.na(.), 0)  # Replace NA values with 0
```


# Step 3: Convert non-numeric columns to numeric as needed
```{r}
numeric_columns <- microbe_data_long %>%
  select(-c(feature, Sample, Site, OTUID)) %>%
  mutate(across(everything(), ~ as.numeric(as.character(.))))
```



# Step 4: Combine the numeric data back with the identifier columns
```{r}
microbe_data_wide <- cbind(microbe_data_long[ , c("feature","Sample", "Site", "OTUID")], numeric_columns)

microbe_data_clean <- na.omit(microbe_data_wide)
```

# Step 5: Create a matrix for dissimilarity calculations
```{r}
microbe_data_matrix <- as.matrix(microbe_data_clean[, -c(1, 2, 3)])
```



# Remove empty columns and rows if any
```{r}
microbe_data_matrix <- microbe_data_matrix[, colSums(microbe_data_matrix != 0) > 0]

microbe_data_matrix <- microbe_data_matrix[rowSums(microbe_data_matrix) > 0, ]
```

# Step 6: Calculate dissimilarity matrices
```{r}
microbe_matrix <- vegdist(microbe_data_matrix, method = "bray")  # Bray-Curtis for microbial data
```

# Step 7: Perform NMDS on microbial data
```{r}
nmds_result <- metaMDS(microbe_matrix, k = 2, trymax = 100)
```

# Step 8: Prepare the NMDS results for plotting
```{r}
nmds_data <- as.data.frame(scores(nmds_result))  # Extract NMDS scores
nmds_data$Sample <- rownames(nmds_data)  # Add sample names
```

# Merge with metadata to include site and depth information
```{r}
nmds_data <- nmds_data %>%
  left_join(microbe_data_long %>% select(Sample, Site, Depth, Abundance), by = "Sample")
```

# Step 9: Plot NMDS results
```{r}
ggplot(nmds_data, aes(x = NMDS1, y = NMDS2, color = Abundance)) +
  geom_point(size = 2, show.legend = TRUE) +
  facet_wrap(~ Depth) +
  theme(text = element_text(size = 24)) +
  xlim(-1.4, 1.7) +
  ylim(-1.51, 1.67) +
  theme_classic() +
  labs(title = "NMDS Plot of Microbial Communities",
       x = "NMDS Axis 1",
       y = "NMDS Axis 2")

```





```{r}
# Load necessary libraries
library(tidyverse)
library(vegan)
library(ggplot2)

# Step 1: Read the data
soil_chem_data <- read.csv("/Users/kyasparks/Desktop/Desktop - MacBook Pro (2)/Kyas_PhD/CSU_2023_2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/Soil_Chemistry_TimFegelRMRS/SoilChemAnalysis_16S_ASCC/ASCC_16S_SoilChemMetadata_rem.csv")

feature <- read.delim("/Users/kyasparks/Desktop/Desktop - MacBook Pro (2)/Kyas_PhD/CSU_2023_2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_metadata/ASCC_16S_feature_table_rem_RELABUND.txt")

# Step 2: Reshape the feature data into a long format
microbe_data_long <- feature %>%
  pivot_longer(cols = 4:306, names_to = "Sample", values_to = "Abundance") %>%
  replace(is.na(.), 0)  # Replace NA values with 0

# Step 3: Convert non-numeric columns to numeric as needed
numeric_columns <- microbe_data_long %>%
  select(-c(feature, Sample, Site, OTUID)) %>%
  mutate(across(everything(), ~ as.numeric(as.character(.))))

# Step 4: Combine the numeric data back with the identifier columns
microbe_data_wide <- cbind(microbe_data_long[ , c("feature","Sample", "Site", "OTUID")], numeric_columns)
microbe_data_clean <- na.omit(microbe_data_wide)

microbe_data_clean <- na.omit(microbe_data_wide)  # Ensure this is applied correctly

# Step 5: Create a matrix for dissimilarity calculations for microbial data
microbe_data_matrix <- as.matrix(microbe_data_clean[, -c(1, 2, 3, 4)])

# Remove empty columns and rows if any
microbe_data_matrix <- microbe_data_matrix[, colSums(microbe_data_matrix != 0) > 0]
microbe_data_matrix <- microbe_data_matrix[rowSums(microbe_data_matrix) > 0, ]

# Check for NAs in the matrix
if (any(is.na(microbe_data_matrix))) {
  stop("NA values found in the microbial data matrix")
}

# Remove empty columns and rows if any
microbe_data_matrix <- microbe_data_matrix[, colSums(microbe_data_matrix != 0) > 0]
microbe_data_matrix <- microbe_data_matrix[rowSums(microbe_data_matrix) > 0, ]

# Step 6: Calculate dissimilarity matrices for microbial data
microbe_matrix <- vegdist(microbe_data_matrix, method = "bray")  # Bray-Curtis for microbial data

# Step 7: Perform PCoA on microbial data
pcoa_microbe <- cmdscale(microbe_matrix, k = 2, eig = TRUE)

# Step 8: Prepare the PCoA results for plotting
pcoa_data <- data.frame(NMDS1 = pcoa_microbe[, 1], NMDS2 = pcoa_microbe[, 2])
pcoa_data$Sample <- rownames(pcoa_data)

# Merge with metadata to include site and depth information
pcoa_data <- left_join(pcoa_data, microbe_data_long %>% select(Sample, Site, Depth), by = "Sample")

# Step 9: Calculate dissimilarity matrices for soil chemistry
soil_data_matrix <- as.matrix(soil_chem_data[, -1])  # Exclude non-numeric columns (like Sample names)
soil_data_matrix <- scale(soil_data_matrix)  # Optionally scale the data
soil_matrix <- dist(soil_data_matrix, method = "euclidean")  # Calculate Euclidean distance for soil chemistry

# Step 10: Fit environmental vectors from soil chemistry data
envfit_results <- envfit(pcoa_microbe, soil_data_matrix, perm = 999)  # Adjust as necessary

# Step 11: Plot PCoA results with environmental vectors
ggplot(pcoa_data, aes(x = NMDS1, y = NMDS2, color = Site)) +
  geom_point(size = 3) +
  geom_segment(data = as.data.frame(scores(envfit_results, "vectors")), 
               aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), 
               arrow = arrow(length = unit(0.2, "inches")), 
               color = "blue") +
  theme_classic() +
  labs(title = "PCoA of Microbial Communities with Soil Chemistry Vectors",
       x = "PCoA Axis 1",
       y = "PCoA Axis 2")
```