---
title: "Cleaning OG MaAsLin Output"
---

## Cleaning output

### Load necessary libraries
```{r}
library(openxlsx)
library(readxl)
library(dplyr)
library(stringr)

# Set file paths
input_file <- "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_metadata/MaaslineResultsCombined_16S_new_042925.xlsx"
output_file <- "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_metadata/MaaslineResultsCombined_16S_new_042925_CLEANED.xlsx"

# Get all sheet names
sheets <- excel_sheets(input_file)

# Create new workbook
wb <- createWorkbook()

# Loop through each sheet, clean, and add to workbook
for (sheet in sheets) {
  df <- read_excel(input_file, sheet = sheet)
  if ("feature" %in% colnames(df)) {
    df <- df %>%
      filter(!str_detect(feature, "\\.\\d$")) %>%
      mutate(feature = feature %>%
               str_replace_all("\\.\\.", "; ") %>%
               str_replace_all("\\.([a-z]__)", "\\1") %>%
               str_remove_all("\\s*;\\s*$") %>%
               str_remove("\\.\\d+$") %>%
               str_trim()
             )
  }
  addWorksheet(wb, sheet)
  writeData(wb, sheet, df)
}

# Save the workbook
saveWorkbook(wb, output_file, overwrite = TRUE)
```

```{r}```{r}
library(openxlsx)
library(readxl)
library(dplyr)
library(stringr)

# File paths
maaslin_file <- "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_metadata/MaaslineResultsCombined_16S_new_042925_CLEANED.xlsx"
output_file <- "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_metadata/MaaslineResultsCombined_16S_new_042925_CLEANED.xlsx"

# Get all sheet names
sheets <- excel_sheets(maaslin_file)

# Create a new workbook
wb <- createWorkbook()

for (sheet in sheets) {
  df <- read_excel(maaslin_file, sheet = sheet)

  # If "feature" exists, rename it to "taxonomy"
  if ("feature" %in% colnames(df)) {
    df <- df %>% rename(taxonomy = feature)
  }

  # NOW: remove rows where taxonomy is exactly "d__Bacteria; ; ; ; ; ;"
  if ("taxonomy" %in% colnames(df)) {
    df <- df %>% filter(taxonomy != "d__Bacteria; ; ; ; ; ;")
  }

  addWorksheet(wb, sheet)
  writeData(wb, sheet, df)
}

# Save the cleaned workbook
saveWorkbook(wb, output_file, overwrite = TRUE)
```


### Merge OTUIDs from taxonomy file into MaAsLin results

```{r}
# File paths
maaslin_file <- "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_metadata/MaaslineResultsCombined_16S_new_042925_CLEANED.xlsx"
taxonomy_file <- "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_taxonomy/ASCC_16S_taxonomy_forprocessing.txt"
output_file <-  "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_metadata/MaaslineResultsCombined_16S_new_042925_CLEANED_withOTUid.xlsx"

# Load taxonomy reference
taxonomy_ref <- read.delim(taxonomy_file, sep = "\t", header = TRUE) %>%  
mutate(taxonomy = str_trim(taxonomy)) %>%  
distinct(taxonomy, .keep_all = TRUE)

# Log sheet names

sheet_names <- excel_sheets(maaslin_file)
print(paste("Sheets to process:", paste(sheet_names, collapse = ", ")))

# Create workbook
wb <- createWorkbook()
for (sheet in sheet_names) {
  message(paste("Processing sheet:", sheet))
  df <- read_excel(maaslin_file, sheet = sheet)
  if (!"taxonomy" %in% colnames(df)) {
    warning(paste("Sheet", sheet, "does not contain 'taxonomy' column. Skipping."))
    next  }
  df <- df %>%    mutate(taxonomy = str_trim(taxonomy)) %>%    inner_join(taxonomy_ref, by = "taxonomy")
  message(paste("Matched rows in", sheet, ":", nrow(df)))
  if (nrow(df) > 0) {
    df <- relocate(df, OTUID, .before = taxonomy)
  }
  addWorksheet(wb, sheet)
  writeData(wb, sheet, df)
}

saveWorkbook(wb, output_file, overwrite = TRUE)
```
