---
title: "ENVFIT"
output: html_notebook
---

```{r}
library(tidyverse)
library(vegan)
library(ggplot2)
library(dplyr)
```


PERMANOVA and NMDS of rarefied table. 

```{r}
metadata <- read.delim("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_metadata/ASCC_16S_metadata_rem.txt",row.names=1) 

# Import feature table
otus <- read.delim("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_metadata/ASCC_16S_feature_table_rem.txt",row.names=1,header=TRUE)
```


We need to invert the dataframe so that each row is a sample and each column is an ASV. This is also an opprotunity to filter out any samples/treatments we are not interested in.
```{r}
#otus.t <- as.data.frame(t(otus[]))

flip <- otus %>% 
 # select(1, !contains("_P1_"))  %>% 
  pivot_longer(1:303, names_to = "Sample", values_to = "count") %>% 
  pivot_wider(names_from = "Sample", values_from = "count")



  # Transpose the data frame
otus_t <- as.data.frame(t(otus))

# Convert row names to a column
otus_t <- tibble::rownames_to_column(otus_t, var = "Samples")


# If you need to pivot the data again, you can use pivot_longer/pivot_wider
# Example of using pivot_longer if needed
otus_long <- otus_t %>%
  pivot_longer(-Sample, names_to = "Sample", values_to = "count")
```

Make sure that the dataframe and metadata have the same number of rows and that the rows are in the exact same order.
```{r}
meta <- otus_t %>% 
  select(,1) %>% 
  inner_join(metadata)
```


Here we filter the dataframe to remove the first column composed of sample names.
```{r}
table <- flip %>% 
  select(2:18550)
```


The next step is to take the data frame and turn it into a distance matrix. Here we are calculating distance using Bray-Curtis distance, but many other methods are available. Distance matrices are pairwise comparisons that compare the similarity of every community on a scale from 0 to 1. 0 indicates communities are exactly the same and 1 indicates there is no similarity in the communities.
```{r}
data.t <- as.matrix(vegdist(table, method = 'bray'))
```

PERMANOVA (Permutational multivariate analysis of variance) is a non-parametric multivariate statisitical test that is useful when studying microbial communities because it allows you to compare how similar different communities are. PERMANOVA is useful because it has very few assumptions that the data need to meet (allows for differences in variation between groups, insensitive to multicollinearity, datasets can be zero-inflated).

Interpreting the output:
F statistic:
R2:
P:
```{r}

mer_perm<-adonis2(data.t ~ Site*Depth, data=meta, permutations = 999)
mer_perm
```

Now we can visualize community similarity using NMDS (non-metric multidimensional scaling), an iterative and non-parametirc method which aims to plot. Additionally, we can compare communities with environmental parameters to understand what enviornmental factors are significantly correlated to community structure using the envfit function from the vegan package. The text file should be the same number of samples and order as the filtered metadata file.


```{r}
NMDS<-metaMDS(data.t, trymax= 200)
envfit.data <- read.delim("october_data/environmental_data.txt") # text file of enviornmental variables
envfit_perm <- envfit(NMDS, envfit.data, permutations = 999) # this fits environmental vectors
```


```{r}
MDS1 = NMDS$points[,1]
MDS2 = NMDS$points[,2]
NMDS = data.frame(NMDS1 = MDS1, NMDS2 = MDS2)
```

```{r}
NMDS_plot <- ggplot(NMDS) +
  geom_point(aes(x=NMDS1, y=NMDS2, col=meta$Site,  shape=meta$Depth), size=4) + 
  theme_bw() +
  labs(color = "Treatment", shape= "Sample Type")+
  #scale_shape_manual(values = c(2, 15)) +  #change shape of points
  #stat_ellipse(aes(x=NMDS1, y=NMDS2, col=Plant)) + #add ellipse around variables
  #scale_color_manual(values=c('darkcyan','goldenrod2','darkmagenta'))+ #change color of points
   theme(text=element_text(size=21))
 NMDS_plot
```

Below is the code to select significant enviornmental variables and overlay the vectors onto the NMDS
```{r}
env.scores <- as.data.frame(scores(envfit_perm, display = "vectors")) #extracts relevant scores from envifit
env.scores <- cbind(env.scores, env.variables = rownames(env.scores)) #and then gives them their names

env.scores <- cbind(env.scores, pval = envfit_perm$vectors$pvals) # add pvalues to dataframe
sig.env.scrs <- subset(env.scores, pval<=0.05) #subset data to show variables significant at 0.05
```


```{r}
plot_env <- NMDS_plot +
  geom_segment(data = sig.env.scrs, aes(x = 0, xend=NMDS1, y=0, yend=NMDS2), arrow = arrow(length = unit(0.25, "cm")), colour = "grey10", lwd=0.3) + #add vector arrows of significant env variables
  ggrepel::geom_text_repel(data = sig.env.scrs, aes(x=NMDS1, y=NMDS2, label = env.variables), cex = 4, direction = "both", segment.size = 0.25)#+ #add labels for env variables
  #labs(title="Ordination with environmental vectors")
plot_env
```

