# ASCC 2023
## Maaslin2 volcano plots redo -> 16S
### 03/05/25
#### Note: Be sure to check the for the discriminant taxa between sites!! Refer to the script located at /Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Rscripts/MaAsLin2/CheckingDiscriminantTaxaByOTU.rmd for additional analysis steps.

**See /Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin2_volcano_plots_wNOTES.md for additional information on analysis steps.**

**Note: This script is for the volcano plots for the Maaslin2 results for the ASCC 2023 16S data.**



**Making volcano plor for SF vs TP shallow**
***I want to overlay the discriminant taxa for all sites within the 0-5cm depth, and I want to do the same with the core taxa across sites & depths***

# Load libraries

```{r}
library(ggplot2)
library(ggrepel)
library(dplyr)
library(readxl)
library(stringr)

# Read in the main dataset (SF v TP shallow)
data <- read_excel("/Users/kyasparks/Desktop/MaaslinResults_includingNonSignif_16S_corrected.xlsx", 
                   sheet = "SF v TP shallow") %>% as.data.frame()

# Read in discriminant results from SF5_discrim (SF-enriched OTUs)

# ✅ Read in discriminant results from SF5_discrim
discriminant_data_sf <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrim_results2.xlsx",
                                   sheet = "SF5_discrim") %>%
  select(OTUID, coef.x, coef.y)

# ✅ Read in discriminant results from TP5_discrim
discriminant_data_tp <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrim_results2.xlsx",
                                   sheet = "TP5_discrim") %>%
  select(OTUID, coef.x, coef.y)

# ✅ Read in core taxa dataset
core_taxa_data <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_CoreTaxaAnalysis/ASCC_16S_CoreTaxa_70perc_allsitescombo.xlsx",
                             sheet = "d70") %>%
  select(OTUID)  # Only keep OTUIDs

# Extract OTUIDs for core taxa
core_taxa <- core_taxa_data$OTUID

# ✅ Identify SF-enriched OTUs (SF higher than TP & SJ)
discriminant_SF <- discriminant_data_sf %>%
  filter(coef.x < 0 & coef.y < 0) %>%   # ✅ FIXED
  select(OTUID) %>%
  pull()

# ✅ Identify TP-enriched OTUs (TP higher than SF & SJ) → **Only appear on the left side**
discriminant_TP <- discriminant_data_tp %>%
  filter(coef.x > 0 & coef.y > 0) %>%   # ✅ FIXED (TP is enriched over SF & SJ)
  select(OTUID) %>%
  pull()

# ✅ Ensure OTUID formatting matches
data$OTUID <- trimws(as.character(data$OTUID))
discriminant_SF <- trimws(as.character(discriminant_SF))
discriminant_TP <- trimws(as.character(discriminant_TP))
core_taxa <- trimws(as.character(core_taxa))

# ✅ Count how many OTUs are now discriminant for SF & TP
num_discriminant_SF <- length(discriminant_SF)
print(paste("Number of SF-discriminant OTUs:", num_discriminant_SF))

num_discriminant_TP <- length(discriminant_TP)
print(paste("Number of TP-discriminant OTUs:", num_discriminant_TP))

# Add columns for visualization
data <- data %>%
  mutate(
    log_q = -log10(qval),
    
    # ✅ Correct enrichment classification
    Enrichment = case_when(
      qval >= 0.05 ~ "Non-Significant",
      coef > 0 ~ "TP Enriched",  
      coef < 0 ~ "SF Enriched"  
    ),
    
    Highlight = case_when(
      OTUID %in% discriminant_SF ~ "Discriminant for SF",
      OTUID %in% discriminant_TP ~ "Discriminant for TP",
      TRUE ~ "Regular"
    ),
    
    Core_Taxa = ifelse(OTUID %in% core_taxa, "Core", NA)
  ) %>%
  mutate(
    Highlight = factor(Highlight, levels = c("Discriminant for SF", "Discriminant for TP", "Regular")),
    Core_Taxa = factor(Core_Taxa, levels = c("Core", NA))
  )


plot <- ggplot(data, aes(x = coef, y = log_q)) +
  
  # All points, with non-significant in grey
  geom_point(aes(color = Enrichment, alpha = Enrichment), size = 3) +
  
  # 🔥 Highlight SF-enriched OTUs (black outline) → Right side
  geom_point(data = data %>% filter(Highlight == "Discriminant for SF" & coef < 0),
             aes(x = coef, y = log_q, fill = Enrichment),
             shape = 21, size = 3, stroke = 0.6, color = "black") +

  geom_point(data = data %>% filter(Highlight == "Discriminant for TP" & coef > 0),
           aes(x = coef, y = log_q, fill = Enrichment),
           shape = 21, size = 3, stroke = 0.6, color = "red") +   # ✅ FIXED COLOR
  
  # 🔴 Highlight TP-enriched OTUs (red outline) → Left side only
  # geom_point(data = data %>% filter(Highlight == "Discriminant for TP" & coef < 0),
  #            aes(x = coef, y = log_q, fill = Enrichment),
  #            shape = 21, size = 3, stroke = 0.6, color = "red") +   # ✅ FIXED COLOR
  
  # 🔵 Circle core taxa (hollow shape)
  geom_point(data = data %>% filter(!is.na(Core_Taxa)),
             aes(x = coef, y = log_q), shape = 21, fill = NA, size = 3, stroke = 1, color = "blue") +

  # ✅ Color scheme: Non-Significant → Grey
  scale_color_manual(values = c(
    "SF Enriched" = "#B75347",  # ✅ Red for SF
    "TP Enriched" = "#94B594",  # ✅ Green for TP
    "Non-Significant" = "#B0B0B0"
  )) +
  
  scale_alpha_manual(values = c("Non-Significant" = 1, "SF Enriched" = 1, "TP Enriched" = 1)) +
  
  scale_fill_manual(values = c(
    "SF Enriched" = "#B75347",
    "TP Enriched" = "#94B594",
    "Non-Significant" = "#B0B0B0"
  )) +
  
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", color = "gray") +
  
  labs(x = "Coefficient", 
       y = "-log10(q-value)", 
       color = "Enrichment", 
       fill = "Enrichment", 
       shape = "Highlighted Taxa") +
  
  theme_minimal() +
  theme(legend.position = "top")

# ✅ Show plot
print(plot)


# Save plot
ggsave(filename = "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Plots/volcano_plot_SF_TP_shallow.svg",
       plot = plot, width = 6, height = 6, dpi = 300, device = "svg")
```


```{r}
# Print OTUIDs of Core Taxa (Circled)
core_otus <- data %>%
  filter(Core_Taxa == "Core") %>%
  select(OTUID)

# View the OTUIDs that were circled
print(core_otus)

# Count how many OTUs were circled
cat("Number of Core Taxa (Circled):", nrow(core_otus), "\n")

# Save the OTUIDs to a CSV file (optional)
write.csv(core_otus, "core_OTUIDssftp5.csv", row.names = FALSE)

# Save to CSV (optional)
write.csv(non_highlighted_otus, "non_highlighted_OTUIDs_sftp5.csv", row.names = FALSE)
```










```{r}
library(ggplot2)
library(ggrepel)
library(dplyr)
library(readxl)
library(stringr)

# Read in the main dataset (SF v TP shallow)
data <- read_excel("/Users/kyasparks/Desktop/MaaslinResults_includingNonSignif_16S_corrected.xlsx", 
                   sheet = "SF v SJ shallow") %>% as.data.frame()

# Read in discriminant results from SF5_discrim (SF-enriched OTUs)

# ✅ Read in discriminant results from SF5_discrim
discriminant_data_sf <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrim_results2.xlsx",
                                   sheet = "SF5_discrim") %>%
  select(OTUID, coef.x, coef.y)

# ✅ Read in discriminant results from TP5_discrim
discriminant_data_sj <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrim_results2.xlsx",
                                   sheet = "SJ5_discrim") %>%
  select(OTUID, coef.x, coef.y)

# ✅ Read in core taxa dataset
core_taxa_data <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_CoreTaxaAnalysis/ASCC_16S_CoreTaxa_70perc_allsitescombo.xlsx",
                             sheet = "d70") %>%
  select(OTUID)  # Only keep OTUIDs

# Extract OTUIDs for core taxa
core_taxa <- core_taxa_data$OTUID

# ✅ Identify SF-enriched OTUs (SF higher than TP & SJ)
discriminant_SF <- discriminant_data_sf %>%
  filter(coef.x < 0 & coef.y < 0) %>%   # ✅ FIXED
  select(OTUID) %>%
  pull()

# ✅ Identify TP-enriched OTUs (TP higher than SF & SJ) → **Only appear on the left side**
discriminant_SJ <- discriminant_data_sj %>%
  filter(coef.x > 0 & coef.y > 0) %>%   # ✅ FIXED (TP is enriched over SF & SJ)
  select(OTUID) %>%
  pull()

# ✅ Ensure OTUID formatting matches
data$OTUID <- trimws(as.character(data$OTUID))
discriminant_SF <- trimws(as.character(discriminant_SF))
discriminant_TP <- trimws(as.character(discriminant_SJ))
core_taxa <- trimws(as.character(core_taxa))

# ✅ Count how many OTUs are now discriminant for SF & TP
num_discriminant_SF <- length(discriminant_SF)
print(paste("Number of SF-discriminant OTUs:", num_discriminant_SF))

num_discriminant_SJ <- length(discriminant_TP)
print(paste("Number of SJ-discriminant OTUs:", num_discriminant_SJ))

# Add columns for visualization
data <- data %>%
  mutate(
    log_q = -log10(qval),
    
    # ✅ Correct enrichment classification
    Enrichment = case_when(
      qval >= 0.05 ~ "Non-Significant",
      coef > 0 ~ "SJ Enriched",  
      coef < 0 ~ "SF Enriched"  
    ),
    
    Highlight = case_when(
      OTUID %in% discriminant_SF ~ "Discriminant for SF",
      OTUID %in% discriminant_SJ ~ "Discriminant for SJ",
      TRUE ~ "Regular"
    ),
    
    Core_Taxa = ifelse(OTUID %in% core_taxa, "Core", NA)
  ) %>%
  mutate(
    Highlight = factor(Highlight, levels = c("Discriminant for SF", "Discriminant for SJ", "Regular")),
    Core_Taxa = factor(Core_Taxa, levels = c("Core", NA))
  )


plot <- ggplot(data, aes(x = coef, y = log_q)) +
  
  # All points, with non-significant in grey
  geom_point(aes(color = Enrichment, alpha = Enrichment), size = 3) +
  
  # 🔥 Highlight SF-enriched OTUs (black outline) → Right side
  geom_point(data = data %>% filter(Highlight == "Discriminant for SF" & coef < 0),
             aes(x = coef, y = log_q, fill = Enrichment),
             shape = 21, size = 3, stroke = 0.6, color = "black") +

  geom_point(data = data %>% filter(Highlight == "Discriminant for SJ" & coef > 0),
           aes(x = coef, y = log_q, fill = Enrichment),
           shape = 21, size = 3, stroke = 0.6, color = "red") +   # ✅ FIXED COLOR
  
  # 🔴 Highlight TP-enriched OTUs (red outline) → Left side only
  # geom_point(data = data %>% filter(Highlight == "Discriminant for TP" & coef < 0),
  #            aes(x = coef, y = log_q, fill = Enrichment),
  #            shape = 21, size = 3, stroke = 0.6, color = "red") +   # ✅ FIXED COLOR
  
  # 🔵 Circle core taxa (hollow shape)
  geom_point(data = data %>% filter(!is.na(Core_Taxa)),
             aes(x = coef, y = log_q), shape = 21, fill = NA, size = 3, stroke = 1, color = "blue") +

  # ✅ Color scheme: Non-Significant → Grey
  scale_color_manual(values = c(
    "SF Enriched" = "#B75347",  # ✅ Red for SF
    "SJ Enriched" = "#E09351",  # ✅ Green for TP
    "Non-Significant" = "#B0B0B0"
  )) +
  
  scale_alpha_manual(values = c("Non-Significant" = 1, "SF Enriched" = 1, "SJ Enriched" = 1)) +
  
  scale_fill_manual(values = c(
    "SF Enriched" = "#B75347",
    "SJ Enriched" ="#E09351",
    "Non-Significant" = "#B0B0B0"
  )) +
  
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", color = "gray") +
  
  labs(x = "Coefficient", 
       y = "-log10(q-value)", 
       color = "Enrichment", 
       fill = "Enrichment", 
       shape = "Highlighted Taxa") +
  
  theme_minimal() +
  theme(legend.position = "top")

# ✅ Show plot
print(plot)


# Save plot
ggsave(filename = "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Plots/volcano_plot_SF_SJ_shallow.svg",
       plot = plot, width = 6, height = 6, dpi = 300, device = "svg")
```


```{r}
# Print OTUIDs of Core Taxa (Circled)
core_otus <- data %>%
  filter(Core_Taxa == "Core") %>%
  select(OTUID)

# View the OTUIDs that were circled
print(core_otus)

# Count how many OTUs were circled
cat("Number of Core Taxa (Circled):", nrow(core_otus), "\n")

# Save the OTUIDs to a CSV file (optional)
write.csv(core_otus, "core_OTUIDssftp5.csv", row.names = FALSE)

# Save to CSV (optional)
write.csv(non_highlighted_otus, "non_highlighted_OTUIDs_sftp5.csv", row.names = FALSE)
```











```{r}
library(ggplot2)
library(ggrepel)
library(dplyr)
library(readxl)
library(stringr)

# Read in the main dataset (SF v TP shallow)
data <- read_excel("/Users/kyasparks/Desktop/MaaslinResults_includingNonSignif_16S_corrected.xlsx", 
                   sheet = "TP v SJ shallow") %>% as.data.frame()

# Read in discriminant results from SF5_discrim (SF-enriched OTUs)

# ✅ Read in discriminant results from SF5_discrim
discriminant_data_sf <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrim_results2.xlsx",
                                   sheet = "TP5_discrim") %>%
  select(OTUID, coef.x, coef.y)

# ✅ Read in discriminant results from TP5_discrim
discriminant_data_sj <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrim_results2.xlsx",
                                   sheet = "SJ5_discrim") %>%
  select(OTUID, coef.x, coef.y)

# ✅ Read in core taxa dataset
core_taxa_data <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_CoreTaxaAnalysis/ASCC_16S_CoreTaxa_70perc_allsitescombo.xlsx",
                             sheet = "d70") %>%
  select(OTUID)  # Only keep OTUIDs

# Extract OTUIDs for core taxa
core_taxa <- core_taxa_data$OTUID

# ✅ Identify SF-enriched OTUs (SF higher than TP & SJ)
discriminant_TP <- discriminant_data_tp %>%
  filter(coef.x > 0 & coef.y > 0) %>%   # ✅ FIXED
  select(OTUID) %>%
  pull()

# ✅ Identify TP-enriched OTUs (TP higher than SF & SJ) → **Only appear on the left side**
discriminant_SJ <- discriminant_data_sj %>%
  filter(coef.x < 0 & coef.y < 0) %>%   # ✅ FIXED (TP is enriched over SF & SJ)
  select(OTUID) %>%
  pull()

# ✅ Ensure OTUID formatting matches
data$OTUID <- trimws(as.character(data$OTUID))
discriminant_SJ <- trimws(as.character(discriminant_SJ))
discriminant_TP <- trimws(as.character(discriminant_TP))
core_taxa <- trimws(as.character(core_taxa))

# ✅ Count how many OTUs are now discriminant for SF & TP
num_discriminant_TP <- length(discriminant_TP)
print(paste("Number of TP-discriminant OTUs:", num_discriminant_TP))

num_discriminant_SJ <- length(discriminant_TP)
print(paste("Number of SJ-discriminant OTUs:", num_discriminant_SJ))

# Add columns for visualization
data <- data %>%
  mutate(
    log_q = -log10(qval),
    
    # ✅ Correct enrichment classification
    Enrichment = case_when(
      qval >= 0.05 ~ "Non-Significant",
      coef < 0 ~ "SJ Enriched",  
      coef > 0 ~ "TP Enriched"  
    ),
    
    Highlight = case_when(
      OTUID %in% discriminant_TP ~ "Discriminant for TP",
      OTUID %in% discriminant_SJ ~ "Discriminant for SJ",
      TRUE ~ "Regular"
    ),
    
    Core_Taxa = ifelse(OTUID %in% core_taxa, "Core", NA)
  ) %>%
  mutate(
    Highlight = factor(Highlight, levels = c("Discriminant for TP", "Discriminant for SJ", "Regular")),
    Core_Taxa = factor(Core_Taxa, levels = c("Core", NA))
  )


plot <- ggplot(data, aes(x = coef, y = log_q)) +
  
  # All points, with non-significant in grey
  geom_point(aes(color = Enrichment, alpha = Enrichment), size = 3) +
  
  # 🔥 Highlight SF-enriched OTUs (black outline) → Right side
  geom_point(data = data %>% filter(Highlight == "Discriminant for TP" & coef >  0),
             aes(x = coef, y = log_q, fill = Enrichment),
             shape = 21, size = 3, stroke = 0.6, color = "black") +

  geom_point(data = data %>% filter(Highlight == "Discriminant for SJ" & coef <  0),
           aes(x = coef, y = log_q, fill = Enrichment),
           shape = 21, size = 3, stroke = 0.6, color = "red") +   # ✅ FIXED COLOR
  
  # 🔴 Highlight TP-enriched OTUs (red outline) → Left side only
  # geom_point(data = data %>% filter(Highlight == "Discriminant for TP" & coef < 0),
  #            aes(x = coef, y = log_q, fill = Enrichment),
  #            shape = 21, size = 3, stroke = 0.6, color = "red") +   # ✅ FIXED COLOR
  
  # 🔵 Circle core taxa (hollow shape)
  geom_point(data = data %>% filter(!is.na(Core_Taxa)),
             aes(x = coef, y = log_q), shape = 21, fill = NA, size = 3, stroke = 1, color = "blue") +

  # ✅ Color scheme: Non-Significant → Grey
  scale_color_manual(values = c(
    "TP Enriched" = "#94B594",  # ✅ Red for SF
    "SJ Enriched" = "#E09351",  # ✅ Green for TP
    "Non-Significant" = "#B0B0B0"
  )) +
  
  scale_alpha_manual(values = c("Non-Significant" = 1, "TP Enriched" = 1, "SJ Enriched" = 1)) +
  
  scale_fill_manual(values = c(
    "TP Enriched" = "#94B594",
    "SJ Enriched" = "#E09351",
    "Non-Significant" = "#B0B0B0"
  )) +
  
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", color = "gray") +
  
  labs(x = "Coefficient", 
       y = "-log10(q-value)", 
       color = "Enrichment", 
       fill = "Enrichment", 
       shape = "Highlighted Taxa") +
  
  theme_minimal() +
  theme(legend.position = "top")

# ✅ Show plot
print(plot)


# Save plot
ggsave(filename = "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Plots/volcano_plot_TP_SJ_shallow.svg",
       plot = plot, width = 6, height = 6, dpi = 300, device = "svg")
```


```{r}
# Print OTUIDs of Core Taxa (Circled)
core_otus <- data %>%
  filter(Core_Taxa == "Core") %>%
  select(OTUID)

# View the OTUIDs that were circled
print(core_otus)

# Count how many OTUs were circled
cat("Number of Core Taxa (Circled):", nrow(core_otus), "\n")

# Save the OTUIDs to a CSV file (optional)
write.csv(core_otus, "core_OTUIDssftp5.csv", row.names = FALSE)

# Save to CSV (optional)
write.csv(non_highlighted_otus, "non_highlighted_OTUIDs_sftp5.csv", row.names = FALSE)
```













----------------------------------------------------------------------------------------------------------------------------


```{r}
library(ggplot2)
library(ggrepel)
library(dplyr)
library(readxl)
library(stringr)

# Read in the main dataset (SF v TP shallow)
data <- read_excel("/Users/kyasparks/Desktop/MaaslinResults_includingNonSignif_16S_corrected.xlsx", 
                   sheet = "SF v TP deep") %>% as.data.frame()

# Read in discriminant results from SF5_discrim (SF-enriched OTUs)

# ✅ Read in discriminant results from SF5_discrim
discriminant_data_sf <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrim_results2.xlsx",
                                   sheet = "SF15_discrim") %>%
  select(OTUID, coef.x, coef.y)

# ✅ Read in discriminant results from TP5_discrim
discriminant_data_tp <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrim_results2.xlsx",
                                   sheet = "TP15_discrim") %>%
  select(OTUID, coef.x, coef.y)

# ✅ Read in core taxa dataset
core_taxa_data <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_CoreTaxaAnalysis/ASCC_16S_CoreTaxa_70perc_allsitescombo.xlsx",
                             sheet = "d70") %>%
  select(OTUID)  # Only keep OTUIDs

# Extract OTUIDs for core taxa
core_taxa <- core_taxa_data$OTUID

# ✅ Identify SF-enriched OTUs (SF higher than TP & SJ)
discriminant_SF <- discriminant_data_sf %>%
  filter(coef.x < 0 & coef.y < 0) %>%   # ✅ FIXED
  select(OTUID) %>%
  pull()

# ✅ Identify TP-enriched OTUs (TP higher than SF & SJ) → **Only appear on the left side**
discriminant_TP <- discriminant_data_tp %>%
  filter(coef.x > 0 & coef.y > 0) %>%   # ✅ FIXED (TP is enriched over SF & SJ)
  select(OTUID) %>%
  pull()

# ✅ Ensure OTUID formatting matches
data$OTUID <- trimws(as.character(data$OTUID))
discriminant_SF <- trimws(as.character(discriminant_SF))
discriminant_TP <- trimws(as.character(discriminant_TP))
core_taxa <- trimws(as.character(core_taxa))

# ✅ Count how many OTUs are now discriminant for SF & TP
num_discriminant_SF <- length(discriminant_SF)
print(paste("Number of SF-discriminant OTUs:", num_discriminant_SF))

num_discriminant_TP <- length(discriminant_TP)
print(paste("Number of TP-discriminant OTUs:", num_discriminant_TP))

# Add columns for visualization
data <- data %>%
  mutate(
    log_q = -log10(qval),
    
    # ✅ Correct enrichment classification
    Enrichment = case_when(
      qval >= 0.05 ~ "Non-Significant",
      coef > 0 ~ "TP Enriched",  
      coef < 0 ~ "SF Enriched"  
    ),
    
    Highlight = case_when(
      OTUID %in% discriminant_SF ~ "Discriminant for SF",
      OTUID %in% discriminant_TP ~ "Discriminant for TP",
      TRUE ~ "Regular"
    ),
    
    Core_Taxa = ifelse(OTUID %in% core_taxa, "Core", NA)
  ) %>%
  mutate(
    Highlight = factor(Highlight, levels = c("Discriminant for SF", "Discriminant for TP", "Regular")),
    Core_Taxa = factor(Core_Taxa, levels = c("Core", NA))
  )


plot <- ggplot(data, aes(x = coef, y = log_q)) +
  
  # All points, with non-significant in grey
  geom_point(aes(color = Enrichment, alpha = Enrichment), size = 3) +
  
  # 🔥 Highlight SF-enriched OTUs (black outline) → Right side
  geom_point(data = data %>% filter(Highlight == "Discriminant for SF" & coef < 0),
             aes(x = coef, y = log_q, fill = Enrichment),
             shape = 21, size = 3, stroke = 0.6, color = "black") +

  geom_point(data = data %>% filter(Highlight == "Discriminant for TP" & coef > 0),
           aes(x = coef, y = log_q, fill = Enrichment),
           shape = 21, size = 3, stroke = 0.6, color = "red") +   # ✅ FIXED COLOR
  
  # 🔴 Highlight TP-enriched OTUs (red outline) → Left side only
  # geom_point(data = data %>% filter(Highlight == "Discriminant for TP" & coef < 0),
  #            aes(x = coef, y = log_q, fill = Enrichment),
  #            shape = 21, size = 3, stroke = 0.6, color = "red") +   # ✅ FIXED COLOR
  
  # 🔵 Circle core taxa (hollow shape)
  geom_point(data = data %>% filter(!is.na(Core_Taxa)),
             aes(x = coef, y = log_q), shape = 21, fill = NA, size = 3, stroke = 1, color = "blue") +

  # ✅ Color scheme: Non-Significant → Grey
  scale_color_manual(values = c(
    "SF Enriched" = "#B75347",  # ✅ Red for SF
    "TP Enriched" = "#94B594",  # ✅ Green for TP
    "Non-Significant" = "#B0B0B0"
  )) +
  
  scale_alpha_manual(values = c("Non-Significant" = 1, "SF Enriched" = 1, "TP Enriched" = 1)) +
  
  scale_fill_manual(values = c(
    "SF Enriched" = "#B75347",
    "TP Enriched" = "#94B594",
    "Non-Significant" = "#B0B0B0"
  )) +
  
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", color = "gray") +
  
  labs(x = "Coefficient", 
       y = "-log10(q-value)", 
       color = "Enrichment", 
       fill = "Enrichment", 
       shape = "Highlighted Taxa") +
  
  theme_minimal() +
  theme(legend.position = "top")

# ✅ Show plot
print(plot)


# Save plot
ggsave(filename = "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Plots/volcano_plot_SF_TP_deep.svg",
       plot = plot, width = 6, height = 6, dpi = 300, device = "svg")
```


```{r}
# Print OTUIDs of Core Taxa (Circled)
core_otus <- data %>%
  filter(Core_Taxa == "Core") %>%
  select(OTUID)

# View the OTUIDs that were circled
print(core_otus)

# Count how many OTUs were circled
cat("Number of Core Taxa (Circled):", nrow(core_otus), "\n")

# Save the OTUIDs to a CSV file (optional)
write.csv(core_otus, "core_OTUIDssftp5.csv", row.names = FALSE)

# Save to CSV (optional)
write.csv(non_highlighted_otus, "non_highlighted_OTUIDs_sftp5.csv", row.names = FALSE)
```










```{r}
library(ggplot2)
library(ggrepel)
library(dplyr)
library(readxl)
library(stringr)

# Read in the main dataset (SF v TP shallow)
data <- read_excel("/Users/kyasparks/Desktop/MaaslinResults_includingNonSignif_16S_corrected.xlsx", 
                   sheet = "SJ v SF deep") %>% as.data.frame()

# Read in discriminant results from SF5_discrim (SF-enriched OTUs)

# ✅ Read in discriminant results from SF5_discrim
discriminant_data_sf <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrim_results2.xlsx",
                                   sheet = "SF15_discrim") %>%
  select(OTUID, coef.x, coef.y)

# ✅ Read in discriminant results from TP5_discrim
discriminant_data_sj <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrim_results2.xlsx",
                                   sheet = "SJ15_discrim") %>%
  select(OTUID, coef.x, coef.y)

# ✅ Read in core taxa dataset
core_taxa_data <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_CoreTaxaAnalysis/ASCC_16S_CoreTaxa_70perc_allsitescombo.xlsx",
                             sheet = "d70") %>%
  select(OTUID)  # Only keep OTUIDs

# Extract OTUIDs for core taxa
core_taxa <- core_taxa_data$OTUID

# ✅ Identify SF-enriched OTUs (SF higher than TP & SJ)
discriminant_SF <- discriminant_data_sf %>%
  filter(coef.x < 0 & coef.y < 0) %>%   # ✅ FIXED
  select(OTUID) %>%
  pull()

# ✅ Identify TP-enriched OTUs (TP higher than SF & SJ) → **Only appear on the left side**
discriminant_SJ <- discriminant_data_sj %>%
  filter(coef.x > 0 & coef.y > 0) %>%   # ✅ FIXED (TP is enriched over SF & SJ)
  select(OTUID) %>%
  pull()

# ✅ Ensure OTUID formatting matches
data$OTUID <- trimws(as.character(data$OTUID))
discriminant_SF <- trimws(as.character(discriminant_SF))
discriminant_TP <- trimws(as.character(discriminant_SJ))
core_taxa <- trimws(as.character(core_taxa))

# ✅ Count how many OTUs are now discriminant for SF & TP
num_discriminant_SF <- length(discriminant_SF)
print(paste("Number of SF-discriminant OTUs:", num_discriminant_SF))

num_discriminant_SJ <- length(discriminant_TP)
print(paste("Number of SJ-discriminant OTUs:", num_discriminant_SJ))

# Add columns for visualization
data <- data %>%
  mutate(
    log_q = -log10(qval),
    
    # ✅ Correct enrichment classification
    Enrichment = case_when(
      qval >= 0.05 ~ "Non-Significant",
      coef > 0 ~ "SJ Enriched",  
      coef < 0 ~ "SF Enriched"  
    ),
    
    Highlight = case_when(
      OTUID %in% discriminant_SF ~ "Discriminant for SF",
      OTUID %in% discriminant_SJ ~ "Discriminant for SJ",
      TRUE ~ "Regular"
    ),
    
    Core_Taxa = ifelse(OTUID %in% core_taxa, "Core", NA)
  ) %>%
  mutate(
    Highlight = factor(Highlight, levels = c("Discriminant for SF", "Discriminant for SJ", "Regular")),
    Core_Taxa = factor(Core_Taxa, levels = c("Core", NA))
  )


plot <- ggplot(data, aes(x = coef, y = log_q)) +
  
  # All points, with non-significant in grey
  geom_point(aes(color = Enrichment, alpha = Enrichment), size = 3) +
  
  # 🔥 Highlight SF-enriched OTUs (black outline) → Right side
  geom_point(data = data %>% filter(Highlight == "Discriminant for SF" & coef < 0),
             aes(x = coef, y = log_q, fill = Enrichment),
             shape = 21, size = 3, stroke = 0.6, color = "black") +

  geom_point(data = data %>% filter(Highlight == "Discriminant for SJ" & coef > 0),
           aes(x = coef, y = log_q, fill = Enrichment),
           shape = 21, size = 3, stroke = 0.6, color = "red") +   # ✅ FIXED COLOR
  
  # 🔴 Highlight TP-enriched OTUs (red outline) → Left side only
  # geom_point(data = data %>% filter(Highlight == "Discriminant for TP" & coef < 0),
  #            aes(x = coef, y = log_q, fill = Enrichment),
  #            shape = 21, size = 3, stroke = 0.6, color = "red") +   # ✅ FIXED COLOR
  
  # 🔵 Circle core taxa (hollow shape)
  geom_point(data = data %>% filter(!is.na(Core_Taxa)),
             aes(x = coef, y = log_q), shape = 21, fill = NA, size = 3, stroke = 1, color = "blue") +

  # ✅ Color scheme: Non-Significant → Grey
  scale_color_manual(values = c(
    "SF Enriched" = "#B75347",  # ✅ Red for SF
    "SJ Enriched" = "#E09351",  # ✅ Green for TP
    "Non-Significant" = "#B0B0B0"
  )) +
  
  scale_alpha_manual(values = c("Non-Significant" = 1, "SF Enriched" = 1, "SJ Enriched" = 1)) +
  
  scale_fill_manual(values = c(
    "SF Enriched" = "#B75347",
    "SJ Enriched" ="#E09351",
    "Non-Significant" = "#B0B0B0"
  )) +
  
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", color = "gray") +
  
  labs(x = "Coefficient", 
       y = "-log10(q-value)", 
       color = "Enrichment", 
       fill = "Enrichment", 
       shape = "Highlighted Taxa") +
  
  theme_minimal() +
  theme(legend.position = "top")

# ✅ Show plot
print(plot)


# Save plot
ggsave(filename = "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Plots/volcano_plot_SF_SJ_DEEP.svg",
       plot = plot, width = 6, height = 6, dpi = 300, device = "svg")
```


```{r}
# Print OTUIDs of Core Taxa (Circled)
core_otus <- data %>%
  filter(Core_Taxa == "Core") %>%
  select(OTUID)

# View the OTUIDs that were circled
print(core_otus)

# Count how many OTUs were circled
cat("Number of Core Taxa (Circled):", nrow(core_otus), "\n")

# Save the OTUIDs to a CSV file (optional)
write.csv(core_otus, "core_OTUIDssftp5.csv", row.names = FALSE)

# Save to CSV (optional)
write.csv(non_highlighted_otus, "non_highlighted_OTUIDs_sftp5.csv", row.names = FALSE)
```











```{r}
library(ggplot2)
library(ggrepel)
library(dplyr)
library(readxl)
library(stringr)

# Read in the main dataset (SF v TP shallow)
data <- read_excel("/Users/kyasparks/Desktop/MaaslinResults_includingNonSignif_16S_corrected.xlsx", 
                   sheet = "TP v SJ deep") %>% as.data.frame()

# Read in discriminant results from SF5_discrim (SF-enriched OTUs)

# ✅ Read in discriminant results from SF5_discrim
discriminant_data_sf <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrim_results2.xlsx",
                                   sheet = "TP15_discrim") %>%
  select(OTUID, coef.x, coef.y)

# ✅ Read in discriminant results from TP5_discrim
discriminant_data_sj <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrim_results2.xlsx",
                                   sheet = "SJ15_discrim") %>%
  select(OTUID, coef.x, coef.y)

# ✅ Read in core taxa dataset
core_taxa_data <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_CoreTaxaAnalysis/ASCC_16S_CoreTaxa_70perc_allsitescombo.xlsx",
                             sheet = "d70") %>%
  select(OTUID)  # Only keep OTUIDs

# Extract OTUIDs for core taxa
core_taxa <- core_taxa_data$OTUID

# ✅ Identify SF-enriched OTUs (SF higher than TP & SJ)
discriminant_TP <- discriminant_data_tp %>%
  filter(coef.x > 0 & coef.y > 0) %>%   # ✅ FIXED
  select(OTUID) %>%
  pull()

# ✅ Identify TP-enriched OTUs (TP higher than SF & SJ) → **Only appear on the left side**
discriminant_SJ <- discriminant_data_sj %>%
  filter(coef.x < 0 & coef.y < 0) %>%   # ✅ FIXED (TP is enriched over SF & SJ)
  select(OTUID) %>%
  pull()

# ✅ Ensure OTUID formatting matches
data$OTUID <- trimws(as.character(data$OTUID))
discriminant_SJ <- trimws(as.character(discriminant_SJ))
discriminant_TP <- trimws(as.character(discriminant_TP))
core_taxa <- trimws(as.character(core_taxa))

# ✅ Count how many OTUs are now discriminant for SF & TP
num_discriminant_TP <- length(discriminant_TP)
print(paste("Number of TP-discriminant OTUs:", num_discriminant_TP))

num_discriminant_SJ <- length(discriminant_TP)
print(paste("Number of SJ-discriminant OTUs:", num_discriminant_SJ))

# Add columns for visualization
data <- data %>%
  mutate(
    log_q = -log10(qval),
    
    # ✅ Correct enrichment classification
    Enrichment = case_when(
      qval >= 0.05 ~ "Non-Significant",
      coef < 0 ~ "SJ Enriched",  
      coef > 0 ~ "TP Enriched"  
    ),
    
    Highlight = case_when(
      OTUID %in% discriminant_TP ~ "Discriminant for TP",
      OTUID %in% discriminant_SJ ~ "Discriminant for SJ",
      TRUE ~ "Regular"
    ),
    
    Core_Taxa = ifelse(OTUID %in% core_taxa, "Core", NA)
  ) %>%
  mutate(
    Highlight = factor(Highlight, levels = c("Discriminant for TP", "Discriminant for SJ", "Regular")),
    Core_Taxa = factor(Core_Taxa, levels = c("Core", NA))
  )


plot <- ggplot(data, aes(x = coef, y = log_q)) +
  
  # All points, with non-significant in grey
  geom_point(aes(color = Enrichment, alpha = Enrichment), size = 3) +
  
  # 🔥 Highlight SF-enriched OTUs (black outline) → Right side
  geom_point(data = data %>% filter(Highlight == "Discriminant for TP" & coef > 0),
             aes(x = coef, y = log_q, fill = Enrichment),
             shape = 21, size = 3, stroke = 0.6, color = "black") +

  geom_point(data = data %>% filter(Highlight == "Discriminant for SJ" & coef < 0),
           aes(x = coef, y = log_q, fill = Enrichment),
           shape = 21, size = 3, stroke = 0.6, color = "red") +   # ✅ FIXED COLOR
  
  # 🔴 Highlight TP-enriched OTUs (red outline) → Left side only
  # geom_point(data = data %>% filter(Highlight == "Discriminant for TP" & coef < 0),
  #            aes(x = coef, y = log_q, fill = Enrichment),
  #            shape = 21, size = 3, stroke = 0.6, color = "red") +   # ✅ FIXED COLOR
  
  # 🔵 Circle core taxa (hollow shape)
  geom_point(data = data %>% filter(!is.na(Core_Taxa)),
             aes(x = coef, y = log_q), shape = 21, fill = NA, size = 3, stroke = 1, color = "blue") +

  # ✅ Color scheme: Non-Significant → Grey
  scale_color_manual(values = c(
    "TP Enriched" = "#94B594",  # ✅ Red for SF
    "SJ Enriched" = "#E09351",  # ✅ Green for TP
    "Non-Significant" = "#B0B0B0"
  )) +
  
  scale_alpha_manual(values = c("Non-Significant" = 1, "TP Enriched" = 1, "SJ Enriched" = 1)) +
  
  scale_fill_manual(values = c(
    "TP Enriched" = "#94B594",
    "SJ Enriched" = "#E09351",
    "Non-Significant" = "#B0B0B0"
  )) +
  
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", color = "gray") +
  
  labs(x = "Coefficient", 
       y = "-log10(q-value)", 
       color = "Enrichment", 
       fill = "Enrichment", 
       shape = "Highlighted Taxa") +
  
  theme_minimal() +
  theme(legend.position = "top")

# ✅ Show plot
print(plot)


# Save plot
ggsave(filename = "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Plots/volcano_plot_TP_SJ_deep.svg",
       plot = plot, width = 6, height = 6, dpi = 300, device = "svg")
```


```{r}
# Print OTUIDs of Core Taxa (Circled)
core_otus <- data %>%
  filter(Core_Taxa == "Core") %>%
  select(OTUID)

# View the OTUIDs that were circled
print(core_otus)

# Count how many OTUs were circled
cat("Number of Core Taxa (Circled):", nrow(core_otus), "\n")

# Save the OTUIDs to a CSV file (optional)
write.csv(core_otus, "core_OTUIDssftp5.csv", row.names = FALSE)

# Save to CSV (optional)
write.csv(non_highlighted_otus, "non_highlighted_OTUIDs_sftp5.csv", row.names = FALSE)
```





















----------------------------------------------------------------------------------------------------------------------------


```{r}
library(ggplot2)
library(ggrepel)
library(dplyr)
library(readxl)
library(stringr)

# Read in the main dataset (SF v TP shallow)
data <- read_excel("/Users/kyasparks/Desktop/MaaslinResults_includingNonSignif_16S_corrected.xlsx", 
                   sheet = "SF v TP OM") %>% as.data.frame()

# Read in discriminant results from SF5_discrim (SF-enriched OTUs)

# ✅ Read in discriminant results from SF5_discrim
discriminant_data_sf <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrim_results2.xlsx",
                                   sheet = "SFOM_discrim") %>%
  select(OTUID, coef.x, coef.y)

# ✅ Read in discriminant results from TP5_discrim
discriminant_data_tp <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrim_results2.xlsx",
                                   sheet = "TPOM_discrim") %>%
  select(OTUID, coef.x, coef.y)

# ✅ Read in core taxa dataset
core_taxa_data <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_CoreTaxaAnalysis/ASCC_16S_CoreTaxa_70perc_allsitescombo.xlsx",
                             sheet = "d70") %>%
  select(OTUID)  # Only keep OTUIDs

# Extract OTUIDs for core taxa
core_taxa <- core_taxa_data$OTUID

# ✅ Identify SF-enriched OTUs (SF higher than TP & SJ)
discriminant_SF <- discriminant_data_sf %>%
  filter(coef.x < 0 & coef.y < 0) %>%   # ✅ FIXED
  select(OTUID) %>%
  pull()

# ✅ Identify TP-enriched OTUs (TP higher than SF & SJ) → **Only appear on the left side**
discriminant_TP <- discriminant_data_tp %>%
  filter(coef.x > 0 & coef.y > 0) %>%   # ✅ FIXED (TP is enriched over SF & SJ)
  select(OTUID) %>%
  pull()

# ✅ Ensure OTUID formatting matches
data$OTUID <- trimws(as.character(data$OTUID))
discriminant_SF <- trimws(as.character(discriminant_SF))
discriminant_TP <- trimws(as.character(discriminant_TP))
core_taxa <- trimws(as.character(core_taxa))

# ✅ Count how many OTUs are now discriminant for SF & TP
num_discriminant_SF <- length(discriminant_SF)
print(paste("Number of SF-discriminant OTUs:", num_discriminant_SF))

num_discriminant_TP <- length(discriminant_TP)
print(paste("Number of TP-discriminant OTUs:", num_discriminant_TP))

# Add columns for visualization
data <- data %>%
  mutate(
    log_q = -log10(qval),
    
    # ✅ Correct enrichment classification
    Enrichment = case_when(
      qval >= 0.05 ~ "Non-Significant",
      coef > 0 ~ "TP Enriched",  
      coef < 0 ~ "SF Enriched"  
    ),
    
    Highlight = case_when(
      OTUID %in% discriminant_SF ~ "Discriminant for SF",
      OTUID %in% discriminant_TP ~ "Discriminant for TP",
      TRUE ~ "Regular"
    ),
    
    Core_Taxa = ifelse(OTUID %in% core_taxa, "Core", NA)
  ) %>%
  mutate(
    Highlight = factor(Highlight, levels = c("Discriminant for SF", "Discriminant for TP", "Regular")),
    Core_Taxa = factor(Core_Taxa, levels = c("Core", NA))
  )


plot <- ggplot(data, aes(x = coef, y = log_q)) +
  
  # All points, with non-significant in grey
  geom_point(aes(color = Enrichment, alpha = Enrichment), size = 3) +
  
  # 🔥 Highlight SF-enriched OTUs (black outline) → Right side
  geom_point(data = data %>% filter(Highlight == "Discriminant for SF" & coef < 0),
             aes(x = coef, y = log_q, fill = Enrichment),
             shape = 21, size = 3, stroke = 0.6, color = "black") +

  geom_point(data = data %>% filter(Highlight == "Discriminant for TP" & coef > 0),
           aes(x = coef, y = log_q, fill = Enrichment),
           shape = 21, size = 3, stroke = 0.6, color = "red") +   # ✅ FIXED COLOR
  
  # 🔴 Highlight TP-enriched OTUs (red outline) → Left side only
  # geom_point(data = data %>% filter(Highlight == "Discriminant for TP" & coef < 0),
  #            aes(x = coef, y = log_q, fill = Enrichment),
  #            shape = 21, size = 3, stroke = 0.6, color = "red") +   # ✅ FIXED COLOR
  
  # 🔵 Circle core taxa (hollow shape)
  geom_point(data = data %>% filter(!is.na(Core_Taxa)),
             aes(x = coef, y = log_q), shape = 21, fill = NA, size = 3, stroke = 1, color = "blue") +

  # ✅ Color scheme: Non-Significant → Grey
  scale_color_manual(values = c(
    "SF Enriched" = "#B75347",  # ✅ Red for SF
    "TP Enriched" = "#94B594",  # ✅ Green for TP
    "Non-Significant" = "#B0B0B0"
  )) +
  
  scale_alpha_manual(values = c("Non-Significant" = 1, "SF Enriched" = 1, "TP Enriched" = 1)) +
  
  scale_fill_manual(values = c(
    "SF Enriched" = "#B75347",
    "TP Enriched" = "#94B594",
    "Non-Significant" = "#B0B0B0"
  )) +
  
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", color = "gray") +
  
  labs(x = "Coefficient", 
       y = "-log10(q-value)", 
       color = "Enrichment", 
       fill = "Enrichment", 
       shape = "Highlighted Taxa") +
  
  theme_minimal() +
  theme(legend.position = "top")

# ✅ Show plot
print(plot)


# Save plot
ggsave(filename = "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Plots/volcano_plot_SF_TP_OM.svg",
       plot = plot, width = 6, height = 6, dpi = 300, device = "svg")
```


```{r}
# Print OTUIDs of Core Taxa (Circled)
core_otus <- data %>%
  filter(Core_Taxa == "Core") %>%
  select(OTUID)

# View the OTUIDs that were circled
print(core_otus)

# Count how many OTUs were circled
cat("Number of Core Taxa (Circled):", nrow(core_otus), "\n")

# Save the OTUIDs to a CSV file (optional)
write.csv(core_otus, "core_OTUIDssftp5.csv", row.names = FALSE)

# Save to CSV (optional)
write.csv(non_highlighted_otus, "non_highlighted_OTUIDs_sftp5.csv", row.names = FALSE)
```










```{r}
library(ggplot2)
library(ggrepel)
library(dplyr)
library(readxl)
library(stringr)

# Read in the main dataset (SF v TP shallow)
data <- read_excel("/Users/kyasparks/Desktop/MaaslinResults_includingNonSignif_16S_corrected.xlsx", 
                   sheet = "SJ v SF OM") %>% as.data.frame()

# Read in discriminant results from SF5_discrim (SF-enriched OTUs)

# ✅ Read in discriminant results from SF5_discrim
discriminant_data_sf <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrim_results2.xlsx",
                                   sheet = "SFOM_discrim") %>%
  select(OTUID, coef.x, coef.y)

# ✅ Read in discriminant results from TP5_discrim
discriminant_data_sj <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrim_results2.xlsx",
                                   sheet = "SJOM_discrim") %>%
  select(OTUID, coef.x, coef.y)

# ✅ Read in core taxa dataset
core_taxa_data <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_CoreTaxaAnalysis/ASCC_16S_CoreTaxa_70perc_allsitescombo.xlsx",
                             sheet = "d70") %>%
  select(OTUID)  # Only keep OTUIDs

# Extract OTUIDs for core taxa
core_taxa <- core_taxa_data$OTUID

# ✅ Identify SF-enriched OTUs (SF higher than TP & SJ)
discriminant_SF <- discriminant_data_sf %>%
  filter(coef.x < 0 & coef.y < 0) %>%   # ✅ FIXED
  select(OTUID) %>%
  pull()

# ✅ Identify TP-enriched OTUs (TP higher than SF & SJ) → **Only appear on the left side**
discriminant_SJ <- discriminant_data_sj %>%
  filter(coef.x > 0 & coef.y > 0) %>%   # ✅ FIXED (TP is enriched over SF & SJ)
  select(OTUID) %>%
  pull()

# ✅ Ensure OTUID formatting matches
data$OTUID <- trimws(as.character(data$OTUID))
discriminant_SF <- trimws(as.character(discriminant_SF))
discriminant_TP <- trimws(as.character(discriminant_SJ))
core_taxa <- trimws(as.character(core_taxa))

# ✅ Count how many OTUs are now discriminant for SF & TP
num_discriminant_SF <- length(discriminant_SF)
print(paste("Number of SF-discriminant OTUs:", num_discriminant_SF))

num_discriminant_SJ <- length(discriminant_TP)
print(paste("Number of SJ-discriminant OTUs:", num_discriminant_SJ))

# Add columns for visualization
data <- data %>%
  mutate(
    log_q = -log10(qval),
    
    # ✅ Correct enrichment classification
    Enrichment = case_when(
      qval >= 0.05 ~ "Non-Significant",
      coef > 0 ~ "SJ Enriched",  
      coef < 0 ~ "SF Enriched"  
    ),
    
    Highlight = case_when(
      OTUID %in% discriminant_SF ~ "Discriminant for SF",
      OTUID %in% discriminant_SJ ~ "Discriminant for SJ",
      TRUE ~ "Regular"
    ),
    
    Core_Taxa = ifelse(OTUID %in% core_taxa, "Core", NA)
  ) %>%
  mutate(
    Highlight = factor(Highlight, levels = c("Discriminant for SF", "Discriminant for SJ", "Regular")),
    Core_Taxa = factor(Core_Taxa, levels = c("Core", NA))
  )


plot <- ggplot(data, aes(x = coef, y = log_q)) +
  
  # All points, with non-significant in grey
  geom_point(aes(color = Enrichment, alpha = Enrichment), size = 3) +
  
  # 🔥 Highlight SF-enriched OTUs (black outline) → Right side
  geom_point(data = data %>% filter(Highlight == "Discriminant for SF" & coef < 0),
             aes(x = coef, y = log_q, fill = Enrichment),
             shape = 21, size = 3, stroke = 0.6, color = "black") +

  geom_point(data = data %>% filter(Highlight == "Discriminant for SJ" & coef > 0),
           aes(x = coef, y = log_q, fill = Enrichment),
           shape = 21, size = 3, stroke = 0.6, color = "red") +   # ✅ FIXED COLOR
  
  # 🔴 Highlight TP-enriched OTUs (red outline) → Left side only
  # geom_point(data = data %>% filter(Highlight == "Discriminant for TP" & coef < 0),
  #            aes(x = coef, y = log_q, fill = Enrichment),
  #            shape = 21, size = 3, stroke = 0.6, color = "red") +   # ✅ FIXED COLOR
  
  # 🔵 Circle core taxa (hollow shape)
  geom_point(data = data %>% filter(!is.na(Core_Taxa)),
             aes(x = coef, y = log_q), shape = 21, fill = NA, size = 3, stroke = 1, color = "blue") +

  # ✅ Color scheme: Non-Significant → Grey
  scale_color_manual(values = c(
    "SF Enriched" = "#B75347",  # ✅ Red for SF
    "SJ Enriched" = "#E09351",  # ✅ Green for TP
    "Non-Significant" = "#B0B0B0"
  )) +
  
  scale_alpha_manual(values = c("Non-Significant" = 1, "SF Enriched" = 1, "SJ Enriched" = 1)) +
  
  scale_fill_manual(values = c(
    "SF Enriched" = "#B75347",
    "SJ Enriched" ="#E09351",
    "Non-Significant" = "#B0B0B0"
  )) +
  
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", color = "gray") +
  
  labs(x = "Coefficient", 
       y = "-log10(q-value)", 
       color = "Enrichment", 
       fill = "Enrichment", 
       shape = "Highlighted Taxa") +
  
  theme_minimal() +
  theme(legend.position = "top")

# ✅ Show plot
print(plot)


# Save plot
ggsave(filename = "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Plots/volcano_plot_SF_SJ_OM.svg",
       plot = plot, width = 6, height = 6, dpi = 300, device = "svg")
```


```{r}
# Print OTUIDs of Core Taxa (Circled)
core_otus <- data %>%
  filter(Core_Taxa == "Core") %>%
  select(OTUID)

# View the OTUIDs that were circled
print(core_otus)

# Count how many OTUs were circled
cat("Number of Core Taxa (Circled):", nrow(core_otus), "\n")

# Save the OTUIDs to a CSV file (optional)
write.csv(core_otus, "core_OTUIDssftp5.csv", row.names = FALSE)

# Save to CSV (optional)
write.csv(non_highlighted_otus, "non_highlighted_OTUIDs_sftp5.csv", row.names = FALSE)
```











```{r}
library(ggplot2)
library(ggrepel)
library(dplyr)
library(readxl)
library(stringr)

# Read in the main dataset (SF v TP shallow)
data <- read_excel("/Users/kyasparks/Desktop/MaaslinResults_includingNonSignif_16S_corrected.xlsx", 
                   sheet = "TP v SJ OM") %>% as.data.frame()

# Read in discriminant results from SF5_discrim (SF-enriched OTUs)

# ✅ Read in discriminant results from SF5_discrim
discriminant_data_sf <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrim_results2.xlsx",
                                   sheet = "TPOM_discrim") %>%
  select(OTUID, coef.x, coef.y)

# ✅ Read in discriminant results from TP5_discrim
discriminant_data_sj <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrim_results2.xlsx",
                                   sheet = "SJOM_discrim") %>%
  select(OTUID, coef.x, coef.y)

# ✅ Read in core taxa dataset
core_taxa_data <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_CoreTaxaAnalysis/ASCC_16S_CoreTaxa_70perc_allsitescombo.xlsx",
                             sheet = "d70") %>%
  select(OTUID)  # Only keep OTUIDs

# Extract OTUIDs for core taxa
core_taxa <- core_taxa_data$OTUID

# ✅ Identify SF-enriched OTUs (SF higher than TP & SJ)
discriminant_TP <- discriminant_data_tp %>%
  filter(coef.x > 0 & coef.y > 0) %>%   # ✅ FIXED
  select(OTUID) %>%
  pull()

# ✅ Identify TP-enriched OTUs (TP higher than SF & SJ) → **Only appear on the left side**
discriminant_SJ <- discriminant_data_sj %>%
  filter(coef.x < 0 & coef.y < 0) %>%   # ✅ FIXED (TP is enriched over SF & SJ)
  select(OTUID) %>%
  pull()

# ✅ Ensure OTUID formatting matches
data$OTUID <- trimws(as.character(data$OTUID))
discriminant_SJ <- trimws(as.character(discriminant_SJ))
discriminant_TP <- trimws(as.character(discriminant_TP))
core_taxa <- trimws(as.character(core_taxa))

# ✅ Count how many OTUs are now discriminant for SF & TP
num_discriminant_TP <- length(discriminant_TP)
print(paste("Number of TP-discriminant OTUs:", num_discriminant_TP))

num_discriminant_SJ <- length(discriminant_TP)
print(paste("Number of SJ-discriminant OTUs:", num_discriminant_SJ))

# Add columns for visualization
data <- data %>%
  mutate(
    log_q = -log10(qval),
    
    # ✅ Correct enrichment classification
    Enrichment = case_when(
      qval >= 0.05 ~ "Non-Significant",
      coef < 0 ~ "SJ Enriched",  
      coef > 0 ~ "TP Enriched"  
    ),
    
    Highlight = case_when(
      OTUID %in% discriminant_TP ~ "Discriminant for TP",
      OTUID %in% discriminant_SJ ~ "Discriminant for SJ",
      TRUE ~ "Regular"
    ),
    
    Core_Taxa = ifelse(OTUID %in% core_taxa, "Core", NA)
  ) %>%
  mutate(
    Highlight = factor(Highlight, levels = c("Discriminant for TP", "Discriminant for SJ", "Regular")),
    Core_Taxa = factor(Core_Taxa, levels = c("Core", NA))
  )


plot <- ggplot(data, aes(x = coef, y = log_q)) +
  
  # All points, with non-significant in grey
  geom_point(aes(color = Enrichment, alpha = Enrichment), size = 3) +
  
  # 🔥 Highlight SF-enriched OTUs (black outline) → Right side
  geom_point(data = data %>% filter(Highlight == "Discriminant for TP" & coef > 0),
             aes(x = coef, y = log_q, fill = Enrichment),
             shape = 21, size = 3, stroke = 0.6, color = "black") +

  geom_point(data = data %>% filter(Highlight == "Discriminant for SJ" & coef < 0),
           aes(x = coef, y = log_q, fill = Enrichment),
           shape = 21, size = 3, stroke = 0.6, color = "red") +   # ✅ FIXED COLOR
  
  # 🔴 Highlight TP-enriched OTUs (red outline) → Left side only
  # geom_point(data = data %>% filter(Highlight == "Discriminant for TP" & coef < 0),
  #            aes(x = coef, y = log_q, fill = Enrichment),
  #            shape = 21, size = 3, stroke = 0.6, color = "red") +   # ✅ FIXED COLOR
  
  # 🔵 Circle core taxa (hollow shape)
  geom_point(data = data %>% filter(!is.na(Core_Taxa)),
             aes(x = coef, y = log_q), shape = 21, fill = NA, size = 3, stroke = 1, color = "blue") +

  # ✅ Color scheme: Non-Significant → Grey
  scale_color_manual(values = c(
    "TP Enriched" = "#94B594",  # ✅ Red for SF
    "SJ Enriched" = "#E09351",  # ✅ Green for TP
    "Non-Significant" = "#B0B0B0"
  )) +
  
  scale_alpha_manual(values = c("Non-Significant" = 1, "TP Enriched" = 1, "SJ Enriched" = 1)) +
  
  scale_fill_manual(values = c(
    "TP Enriched" = "#94B594",
    "SJ Enriched" = "#E09351",
    "Non-Significant" = "#B0B0B0"
  )) +
  
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", color = "gray") +
  
  labs(x = "Coefficient", 
       y = "-log10(q-value)", 
       color = "Enrichment", 
       fill = "Enrichment", 
       shape = "Highlighted Taxa") +
  
  theme_minimal() +
  theme(legend.position = "top")

# ✅ Show plot
print(plot)


# Save plot
ggsave(filename = "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Plots/volcano_plot_TP_SJ_OM.svg",
       plot = plot, width = 6, height = 6, dpi = 300, device = "svg")
```


```{r}
# Print OTUIDs of Core Taxa (Circled)
core_otus <- data %>%
  filter(Core_Taxa == "Core") %>%
  select(OTUID)

# View the OTUIDs that were circled
print(core_otus)

# Count how many OTUs were circled
cat("Number of Core Taxa (Circled):", nrow(core_otus), "\n")

# Save the OTUIDs to a CSV file (optional)
write.csv(core_otus, "core_OTUIDssftp5.csv", row.names = FALSE)

# Save to CSV (optional)
write.csv(non_highlighted_otus, "non_highlighted_OTUIDs_sftp5.csv", row.names = FALSE)
```



