---
title: "Updated Volcano Plots for MaAsLin2 Results 4/8/2024: 16S & ITS"
---

# ASCC 2023
## This script is used to plot volcano plots for the MaAsLin2 results
### Plotting volcano plots, highlighting consistent taxa across sites, and highlight outliers.
#### Note: Be sure to check the for the discriminant taxa between sites!! Refer to the script located at /Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Rscripts/MaAsLin2/CheckingDiscriminantTaxaByOTU_040825.rmd

- sf v. sj shallow: when the coef is negative it is discriminant for SF and when its positive its discriminant for SJ
- sf v. tp shallow: when the coef is negative it is discriminant for SF and when its positive its discriminant for TP
- tp v. sj shallow: when the coef is negative it is discriminant for SJ and when its positive its discriminant for TP
- sj v. sf deep: when the coef is negative it is discriminant for SF and when its positive its discriminant for SJ
- sf v. tp deep: when the coef is negative it is discriminant for SF and when its positive its discriminant for TP
- tp v. sj deep: when the coef is negative it is discriminant for SJ and when its positive its discriminant for TP
- sj v. sf om: when the coef is negative it is discriminant for SF and when its positive its discriminant for SJ
- sf v. tp om: when the coef is negative it is discriminant for SF and when its positive its discriminant for TP
- tp v. sj om: when the coef is negative it is discriminant for SJ and when its positive its discriminant for TP
the colors im using are

# 16S

```{r}
library(ggplot2)
library(ggrepel)
library(dplyr)
library(readxl)
library(stringr)
```

# SF v TP shallow 0-5 cm
```{r}
data <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_allresults_ComBINED_16S.xlsx", sheet = "SF v TP shallow") %>% 
  mutate(OTUID = trimws(as.character(OTUID)))

data <- data %>%
  mutate(
    tax_parts = str_split(taxonomy, ";\\s*"),
    family = sapply(tax_parts, function(x) if (length(x) > 0) x[grepl("^f__", x)][1] else NA)
  )

# Load discriminant OTUs
sf5_discrim <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrimcomparisons_040725.xlsx",sheet = "SF5_discrim") %>% 
  pull(OTUID) %>% trimws()

tp5_discrim <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrimcomparisons_040725.xlsx", sheet = "TP5_discrim") %>% 
  pull(OTUID) %>% trimws()

# Process data
data <- data %>%
  mutate(
    log_q = -log10(qval),
    Enrichment = case_when(
      qval >= 0.05 ~ "Non-Significant",
      coef < 0 ~ "SF Enriched",
      coef > 0 ~ "TP Enriched"
    ),
    Discriminant = OTUID %in% c(sf5_discrim, tp5_discrim),
    tax_parts = str_split(taxonomy, ";\\s*"),
    class = sapply(tax_parts, function(x) x[grepl("^c__", x)][1]),
    order = sapply(tax_parts, function(x) x[grepl("^o__", x)][1]),
    family = sapply(tax_parts, function(x) x[grepl("^f__", x)][1]),
    genus  = sapply(tax_parts, function(x) x[grepl("^g__", x)][1]),
    species = sapply(tax_parts, function(x) x[grepl("^s__", x)][1]),
    LabelText = family, # Use only family names for labels
    Label = ifelse(log_q > 7, LabelText, NA) # Apply threshold for labeling
  )


# Plot
# Plot
plot_sf_tp_shallow <- ggplot(data, aes(x = coef, y = log_q)) +
  geom_point(aes(fill = Enrichment),
             shape = 21, size = 3, stroke = 0, color = "black", alpha = 0.7) +
  geom_point(data = filter(data, Discriminant),
             aes(fill = Enrichment),
             shape = 21, size = 3, stroke = 0.6, color = "black", alpha = 0.7) +
  geom_text_repel(data = filter(data, !is.na(Label)),
                  aes(label = Label),
                  size = 2.5, max.overlaps = 25, segment.color = "gray50") +
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", color = "gray50") + # Add threshold line
  geom_vline(xintercept = 0, linetype = "dashed") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) + # Ensure y-axis includes 0
  scale_fill_manual(values = c(
    "SF Enriched" = "#B75347",
    "TP Enriched" = "#94B594",
    "Non-Significant" = "gray80"
  )) +
  labs(x = "Coefficient", y = "-log10(q-value)", fill = "Enrichment") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(), # Remove gridlines
    axis.line = element_line(color = "black"), # Add black axis lines
    axis.ticks = element_line(color = "black"), # Add black axis ticks
    axis.text = element_text(size = 12, color = "black"), # Customize axis text
    axis.title = element_text(size = 14, color = "black") # Customize axis titles
  )

print(plot_sf_tp_shallow)

# Save plot
ggsave(filename = "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Plots/volcano_SF_v_TP_shallow.svg",
       plot = plot_sf_tp_shallow, width = 6, height = 6, dpi = 300, device = "svg")

# Save outliers
outliers_df <- data %>%
  filter(log_q > 7) %>%
  select(OTUID, taxonomy, LabelText, coef, qval, log_q, Enrichment)

write.csv(outliers_df,
          file = "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Plots/outliers_SF_v_TP_shallow.csv",
          row.names = FALSE)
```



# SJ v TP shallow 0-5 cm

```{r}
# Load main results
data <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_allresults_ComBINED_16S.xlsx", sheet = "TP v SJ shallow") %>% 
  mutate(OTUID = trimws(as.character(OTUID)))

# Load discriminant OTUs
sj5_discrim <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrimcomparisons_040725.xlsx", sheet = "SJ5_discrim") %>% 
  pull(OTUID) %>% trimws()

tp5_discrim <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrimcomparisons_040725.xlsx", sheet = "TP5_discrim") %>% 
  pull(OTUID) %>% trimws()

# Process data
data <- data %>%
  mutate(
    log_q = -log10(qval),
    Enrichment = case_when(
      qval >= 0.05 ~ "Non-Significant",
      coef < 0 ~ "SJ Enriched",
      coef > 0 ~ "TP Enriched"
    ),
    Discriminant = OTUID %in% c(sj5_discrim, tp5_discrim),
    tax_parts = str_split(taxonomy, ";\\s*"),
    family = sapply(tax_parts, function(x) x[grepl("^f__", x)][1]), # Extract family names
    LabelText = family, # Use only family names for labels
    Label = ifelse(log_q > 7, LabelText, NA) # Apply threshold for labeling
  )

# Plot
plot_tp_sj_shallow <- ggplot(data, aes(x = coef, y = log_q)) +
  geom_point(aes(fill = Enrichment),
             shape = 21, size = 3, stroke = 0, color = "black", alpha = 0.7) +
  geom_point(data = filter(data, Discriminant),
             aes(fill = Enrichment),
             shape = 21, size = 3, stroke = 0.6, color = "black", alpha = 0.7) +
  geom_text_repel(data = filter(data, !is.na(Label)),
                  aes(label = Label),
                  size = 2.5, max.overlaps = 25, segment.color = "gray50") +
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", color = "gray50") +
  geom_hline(yintercept = 0, linetype = "solid", color = "black") + # Ensure x-axis crosses at y=0
  geom_vline(xintercept = 0, linetype = "dashed") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) + # Ensure y-axis includes 0
  scale_fill_manual(values = c(
    "SJ Enriched" = "#E09351",
    "TP Enriched" = "#94B594",
    "Non-Significant" = "gray80"
  )) +
  labs(x = "Coefficient", y = "-log10(q-value)", fill = "Enrichment") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(), # Remove gridlines
    axis.line = element_line(color = "black"), # Add black axis lines
    axis.ticks = element_line(color = "black"), # Add black axis ticks
    axis.text = element_text(size = 12, color = "black"), # Customize axis text
    axis.title = element_text(size = 14, color = "black") # Customize axis titles
  )

print(plot_tp_sj_shallow)

# Save plot
ggsave(filename = "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Plots/volcano_TP_v_SJ_shallow.svg",
       plot = (plot_tp_sj_shallow, width = 6, height = 6, dpi = 300, device = "svg"))

# Save outliers
outliers_df <- data %>%
  filter(log_q > 7) %>%
  select(OTUID, taxonomy, LabelText, coef, qval, log_q, Enrichment)

write.csv(outliers_df,
          file = "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Plots/outliers_TP_v_SJ_shallow.csv",
          row.names = FALSE)
```





# SJ v SF shallow 0-5 cm

```{r}
# Load main results
data <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_allresults_ComBINED_16S.xlsx", sheet = "SJ v SF shallow") %>% 
  mutate(OTUID = trimws(as.character(OTUID)))

# Load discriminant OTUs
sj5_discrim <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrimcomparisons_040725.xlsx", sheet = "SJ5_discrim") %>% 
  pull(OTUID) %>% trimws()

sf5_discrim <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrimcomparisons_040725.xlsx", sheet = "SF5_discrim") %>% 
  pull(OTUID) %>% trimws()

# Process data
data <- data %>%
  mutate(
    log_q = -log10(qval),
    Enrichment = case_when(
      qval >= 0.05 ~ "Non-Significant",
      coef < 0 ~ "SF Enriched",
      coef > 0 ~ "SJ Enriched"
    ),
    Discriminant = OTUID %in% c(sf5_discrim, sj5_discrim),
    tax_parts = str_split(taxonomy, ";\\s*"),
    family = sapply(tax_parts, function(x) x[grepl("^f__", x)][1]), # Extract family names
    LabelText = family, # Use only family names for labels
    Label = ifelse(log_q > 7, LabelText, NA) # Apply threshold for labeling
  )

# Plot
plot_sf_sj_shallow <- ggplot(data, aes(x = coef, y = log_q)) +
  geom_point(aes(fill = Enrichment),
             shape = 21, size = 3, stroke = 0, color = "black", alpha = 0.7) +
  geom_point(data = filter(data, Discriminant),
             aes(fill = Enrichment),
             shape = 21, size = 3, stroke = 0.6, color = "black", alpha = 0.7) +
  geom_text_repel(data = filter(data, !is.na(Label)),
                  aes(label = Label),
                  size = 2.5, max.overlaps = 25, segment.color = "gray50") +
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", color = "gray50") +
  geom_hline(yintercept = 0, linetype = "solid", color = "black") + # Ensure x-axis crosses at y=0
  geom_vline(xintercept = 0, linetype = "dashed") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) + # Ensure y-axis includes 0
  scale_fill_manual(values = c(
    "SF Enriched" = "#B75347",
    "SJ Enriched" = "#E09351",
    "Non-Significant" = "gray80"
  )) +
  labs(x = "Coefficient", y = "-log10(q-value)", fill = "Enrichment") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(), # Remove gridlines
    axis.line = element_line(color = "black"), # Add black axis lines
    axis.ticks = element_line(color = "black"), # Add black axis ticks
    axis.text = element_text(size = 12, color = "black"), # Customize axis text
    axis.title = element_text(size = 14, color = "black") # Customize axis titles
  )

print(plot_sf_sj_shallow)

# Save plot
ggsave(filename = "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Plots/volcano_SJ_v_SF_shallow.svg",
       plot = plot_sf_sj_shallow, width = 6, height = 6, dpi = 300, device = "svg")

# Save outliers
outliers_df <- data %>%
  filter(log_q > 7) %>%
  select(OTUID, taxonomy, LabelText, coef, qval, log_q, Enrichment)

write.csv(outliers_df,
          file = "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Plots/outliers_SJ_v_SF_shallow.csv",
          row.names = FALSE)
```


# SJ v SF deep 5-15 cm
```{r}
# Load main results
data <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_allresults_ComBINED_16S.xlsx", sheet = "SJ v SF deep") %>% 
  mutate(OTUID = trimws(as.character(OTUID)))

# Load discriminant OTUs
sj15_discrim <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrimcomparisons_040725.xlsx", sheet = "SJ15_discrim") %>% 
  pull(OTUID) %>% trimws()

sf15_discrim <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrimcomparisons_040725.xlsx", sheet = "SF15_discrim") %>% 
  pull(OTUID) %>% trimws()

# Process data
data <- data %>%
  mutate(
    log_q = -log10(qval),
    Enrichment = case_when(
      qval >= 0.05 ~ "Non-Significant",
      coef < 0 ~ "SF Enriched",
      coef > 0 ~ "SJ Enriched"
    ),
    Discriminant = OTUID %in% c(sf15_discrim, sj15_discrim),
    tax_parts = str_split(taxonomy, ";\\s*"),
    family = sapply(tax_parts, function(x) x[grepl("^f__", x)][1]), # Extract family names
    LabelText = family, # Use only family names for labels
    Label = ifelse(log_q > 10, LabelText, NA) # Apply threshold for labeling
  )

# Plot
plot_sf_sj_deep <- ggplot(data, aes(x = coef, y = log_q)) +
  geom_point(aes(fill = Enrichment),
             shape = 21, size = 3, stroke = 0, color = "black", alpha = 0.7) +
  geom_point(data = filter(data, Discriminant),
             aes(fill = Enrichment),
             shape = 21, size = 3, stroke = 0.6, color = "black", alpha = 0.7) +
  geom_text_repel(data = filter(data, !is.na(Label)),
                  aes(label = Label),
                  size = 2.5, max.overlaps = 25, segment.color = "gray50") +
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", color = "gray50") +
  geom_hline(yintercept = 0, linetype = "solid", color = "black") + # Ensure x-axis crosses at y=0
  geom_vline(xintercept = 0, linetype = "dashed") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) + # Ensure y-axis includes 0
  scale_fill_manual(values = c(
    "SF Enriched" = "#B75347",
    "SJ Enriched" = "#E09351",
    "Non-Significant" = "gray80"
  )) +
  labs(x = "Coefficient", y = "-log10(q-value)", fill = "Enrichment") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(), # Remove gridlines
    axis.line = element_line(color = "black"), # Add black axis lines
    axis.ticks = element_line(color = "black"), # Add black axis ticks
    axis.text = element_text(size = 12, color = "black"), # Customize axis text
    axis.title = element_text(size = 14, color = "black") # Customize axis titles
  )

print(plot_sf_sj_deep)

# Save plot
ggsave(filename = "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Plots/volcano_SJ_v_SF_deep.svg",
       plot = plot_sf_sj_deep, width = 6, height = 6, dpi = 300, device = "svg")

# Save outliers
outliers_df <- data %>%
  filter(log_q > 10) %>%
  select(OTUID, taxonomy, LabelText, coef, qval, log_q, Enrichment)

write.csv(outliers_df,
          file = "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Plots/outliers_SJ_v_SF_deep.csv",
          row.names = FALSE)
```


# SF v TP deep 5-15 cm
```{r}
# Load main results
data <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_allresults_ComBINED_16S.xlsx", sheet = "SF v TP deep") %>% 
  mutate(OTUID = trimws(as.character(OTUID)))

# Load discriminant OTUs
sf15_discrim <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrimcomparisons_040725.xlsx", sheet = "SF15_discrim") %>% 
  pull(OTUID) %>% trimws()

tp15_discrim <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrimcomparisons_040725.xlsx", sheet = "TP15_discrim") %>% 
  pull(OTUID) %>% trimws()

# Process data
data <- data %>%
  mutate(
    log_q = -log10(qval),
    Enrichment = case_when(
      qval >= 0.05 ~ "Non-Significant",
      coef < 0 ~ "SF Enriched",
      coef > 0 ~ "TP Enriched"
    ),
    Discriminant = OTUID %in% c(sf15_discrim, tp15_discrim),
    tax_parts = str_split(taxonomy, ";\\s*"),
    family = sapply(tax_parts, function(x) x[grepl("^f__", x)][1]), # Extract family names
    LabelText = family, # Use only family names for labels
    Label = ifelse(log_q > 2.75, LabelText, NA) # Apply threshold for labeling
  )

# Plot
plot_sf_tp_deep <- ggplot(data, aes(x = coef, y = log_q)) +
  geom_point(aes(fill = Enrichment),
             shape = 21, size = 3, stroke = 0, color = "black", alpha = 0.7) +
  geom_point(data = filter(data, Discriminant),
             aes(fill = Enrichment),
             shape = 21, size = 3, stroke = 0.6, color = "black", alpha = 0.7) +
  geom_text_repel(data = filter(data, !is.na(Label)),
                  aes(label = Label),
                  size = 2.5, max.overlaps = 25, segment.color = "gray50") +
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", color = "gray50") +
  geom_hline(yintercept = 0, linetype = "solid", color = "black") + # Ensure x-axis crosses at y=0
  geom_vline(xintercept = 0, linetype = "dashed") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) + # Ensure y-axis includes 0
  scale_fill_manual(values = c(
    "SF Enriched" = "#B75347",
    "TP Enriched" = "#94B594",
    "Non-Significant" = "gray80"
  )) +
  labs(x = "Coefficient", y = "-log10(q-value)", fill = "Enrichment") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(), # Remove gridlines
    axis.line = element_line(color = "black"), # Add black axis lines
    axis.ticks = element_line(color = "black"), # Add black axis ticks
    axis.text = element_text(size = 12, color = "black"), # Customize axis text
    axis.title = element_text(size = 14, color = "black") # Customize axis titles
  )

print(plot_sf_tp_deep)

# Save plot
ggsave(filename = "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Plots/volcano_SF_v_TP_deep.svg",
       plot = plot_sf_tp_deep, width = 6, height = 6, dpi = 300, device = "svg")

# Save outliers
outliers_df <- data %>%
  filter(log_q > 2.75) %>%
  select(OTUID, taxonomy, LabelText, coef, qval, log_q, Enrichment)

write.csv(outliers_df,
          file = "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Plots/outliers_SF_v_TP_deep.csv",
          row.names = FALSE)
```



# TP v SJ deep 5-15 cm
```{r}
# Load main results
data <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_allresults_ComBINED_16S.xlsx", sheet = "TP v SJ deep") %>% 
  mutate(OTUID = trimws(as.character(OTUID)))

# Load discriminant OTUs (15 cm)
tp15_discrim <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrimcomparisons_040725.xlsx", sheet = "TP15_discrim") %>% 
  pull(OTUID) %>% trimws()

sj15_discrim <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrimcomparisons_040725.xlsx", sheet = "SJ15_discrim") %>% 
  pull(OTUID) %>% trimws()

# Process data
data <- data %>%
  mutate(
    log_q = -log10(qval),
    Enrichment = case_when(
      qval >= 0.05 ~ "Non-Significant",
      coef < 0 ~ "SJ Enriched",
      coef > 0 ~ "TP Enriched"
    ),
    Discriminant = OTUID %in% c(tp15_discrim, sj15_discrim),
    tax_parts = str_split(taxonomy, ";\\s*"),
    family = sapply(tax_parts, function(x) x[grepl("^f__", x)][1]), # Extract family names
    LabelText = family, # Use only family names for labels
    Label = ifelse(log_q > 8, LabelText, NA) # Apply threshold for labeling
  )

# Plot
plot_sj_tp_deep <- ggplot(data, aes(x = coef, y = log_q)) +
  geom_point(aes(fill = Enrichment),
             shape = 21, size = 3, stroke = 0, color = "black", alpha = 0.7) +
  geom_point(data = filter(data, Discriminant),
             aes(fill = Enrichment),
             shape = 21, size = 3, stroke = 0.6, color = "black", alpha = 0.7) +
  geom_text_repel(data = filter(data, !is.na(Label)),
                  aes(label = Label),
                  size = 2.5, max.overlaps = 25, segment.color = "gray50") +
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", color = "gray50") +
  geom_hline(yintercept = 0, linetype = "solid", color = "black") + # Ensure x-axis crosses at y=0
  geom_vline(xintercept = 0, linetype = "dashed") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) + # Ensure y-axis includes 0
  scale_fill_manual(values = c(
    "SJ Enriched" = "#E09351",
    "TP Enriched" = "#94B594",
    "Non-Significant" = "gray80"
  )) +
  labs(x = "Coefficient", y = "-log10(q-value)", fill = "Enrichment") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(), # Remove gridlines
    axis.line = element_line(color = "black"), # Add black axis lines
    axis.ticks = element_line(color = "black"), # Add black axis ticks
    axis.text = element_text(size = 12, color = "black"), # Customize axis text
    axis.title = element_text(size = 14, color = "black") # Customize axis titles
  )

print(plot_sj_tp_deep)

# Save plot
ggsave(filename = "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Plots/volcano_TP_v_SJ_deep.svg",
       plot = plot_sj_tp_deep, width = 6, height = 6, dpi = 300, device = "svg")

# Save outliers
outliers_df <- data %>%
  filter(log_q > 8) %>%
  select(OTUID, taxonomy, LabelText, coef, qval, log_q, Enrichment)

write.csv(outliers_df,
          file = "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Plots/outliers_TP_v_SJ_deep.csv",
          row.names = FALSE)
```




# SJ v SF om
```{r}
# Load main results
data <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_allresults_ComBINED_16S.xlsx", sheet = "SJ v SF OM") %>% 
  mutate(OTUID = trimws(as.character(OTUID)))

# Load discriminant OTUs
sjom_discrim <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrimcomparisons_040725.xlsx", sheet = "SJOM_discrim") %>% 
  pull(OTUID) %>% trimws()

sfom_discrim <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrimcomparisons_040725.xlsx", sheet = "SFOM_discrim") %>% 
  pull(OTUID) %>% trimws()

# Process data
data <- data %>%
  mutate(
    log_q = -log10(qval),
    Enrichment = case_when(
      qval >= 0.05 ~ "Non-Significant",
      coef < 0 ~ "SF Enriched",
      coef > 0 ~ "SJ Enriched"
    ),
    Discriminant = OTUID %in% c(sfom_discrim, sjom_discrim),
    tax_parts = str_split(taxonomy, ";\\s*"),
    family = sapply(tax_parts, function(x) x[grepl("^f__", x)][1]), # Extract family names
    LabelText = family, # Use only family names for labels
    Label = ifelse(log_q > 5, LabelText, NA) # Apply threshold for labeling
  )

# Plot
plot_sj_sf_om <- ggplot(data, aes(x = coef, y = log_q)) +
  geom_point(aes(fill = Enrichment),
             shape = 21, size = 3, stroke = 0, color = "black", alpha = 0.7) +
  geom_point(data = filter(data, Discriminant),
             aes(fill = Enrichment),
             shape = 21, size = 3, stroke = 0.6, color = "black", alpha = 0.7) +
  geom_text_repel(data = filter(data, !is.na(Label)),
                  aes(label = Label),
                  size = 2.5, max.overlaps = 25, segment.color = "gray50") +
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", color = "gray50") +
  geom_hline(yintercept = 0, linetype = "solid", color = "black") + # Ensure x-axis crosses at y=0
  geom_vline(xintercept = 0, linetype = "dashed") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) + # Ensure y-axis includes 0
  scale_fill_manual(values = c(
    "SF Enriched" = "#B75347",
    "SJ Enriched" = "#E09351",
    "Non-Significant" = "gray80"
  )) +
  labs(x = "Coefficient", y = "-log10(q-value)", fill = "Enrichment") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(), # Remove gridlines
    axis.line = element_line(color = "black"), # Add black axis lines
    axis.ticks = element_line(color = "black"), # Add black axis ticks
    axis.text = element_text(size = 12, color = "black"), # Customize axis text
    axis.title = element_text(size = 14, color = "black") # Customize axis titles
  )

print(plot_sj_sf_om)

# Save plot
ggsave(filename = "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Plots/volcano_SJ_v_SF_om.svg",
       plot = plot_sj_sf_om, width = 6, height = 6, dpi = 300, device = "svg")

# Save outliers
outliers_df <- data %>%
  filter(log_q > 5) %>%
  select(OTUID, taxonomy, LabelText, coef, qval, log_q, Enrichment)

write.csv(outliers_df,
          file = "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Plots/outliers_SJ_v_SF_om.csv",
          row.names = FALSE)
```

# SF v TP om
```{r}
# Load main results
data <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_allresults_ComBINED_16S.xlsx", sheet = "SF v TP OM") %>% 
  mutate(OTUID = trimws(as.character(OTUID)))

# Load discriminant OTUs
sfom_discrim <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrimcomparisons_040725.xlsx", sheet = "SFOM_discrim") %>% 
  pull(OTUID) %>% trimws()

tpom_discrim <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrimcomparisons_040725.xlsx", sheet = "TPOM_discrim") %>% 
  pull(OTUID) %>% trimws()

# Process data
data <- data %>%
  mutate(
    log_q = -log10(qval),
    Enrichment = case_when(
      qval >= 0.05 ~ "Non-Significant",
      coef < 0 ~ "SF Enriched",
      coef > 0 ~ "TP Enriched"
    ),
    Discriminant = OTUID %in% c(sfom_discrim, tpom_discrim),
    tax_parts = str_split(taxonomy, ";\\s*"),
    family = sapply(tax_parts, function(x) x[grepl("^f__", x)][1]), # Extract family names
    LabelText = family, # Use only family names for labels
    Label = ifelse(log_q > 6, LabelText, NA) # Apply threshold for labeling
  )

# Plot
plot_tp_sf_om <- ggplot(data, aes(x = coef, y = log_q)) +
  geom_point(aes(fill = Enrichment),
             shape = 21, size = 3, stroke = 0, color = "black", alpha = 0.7) +
  geom_point(data = filter(data, Discriminant),
             aes(fill = Enrichment),
             shape = 21, size = 3, stroke = 0.6, color = "black", alpha = 0.7) +
  geom_text_repel(data = filter(data, !is.na(Label)),
                  aes(label = Label),
                  size = 2.5, max.overlaps = 25, segment.color = "gray50") +
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", color = "gray50") +
  geom_hline(yintercept = 0, linetype = "solid", color = "black") + # Ensure x-axis crosses at y=0
  geom_vline(xintercept = 0, linetype = "dashed") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) + # Ensure y-axis includes 0
  scale_fill_manual(values = c(
    "SF Enriched" = "#B75347",
    "TP Enriched" = "#94B594",
    "Non-Significant" = "gray80"
  )) +
  labs(x = "Coefficient", y = "-log10(q-value)", fill = "Enrichment") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(), # Remove gridlines
    axis.line = element_line(color = "black"), # Add black axis lines
    axis.ticks = element_line(color = "black"), # Add black axis ticks
    axis.text = element_text(size = 12, color = "black"), # Customize axis text
    axis.title = element_text(size = 14, color = "black") # Customize axis titles
  )

print(plot_tp_sf_om)

# Save plot
ggsave(filename = "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Plots/volcano_SF_v_TP_om.svg",
       plot = plot_tp_sf_om, width = 6, height = 6, dpi = 300, device = "svg")

# Save outliers
outliers_df <- data %>%
  filter(log_q > 6) %>%
  select(OTUID, taxonomy, LabelText, coef, qval, log_q, Enrichment)

write.csv(outliers_df,
          file = "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Plots/outliers_SF_v_TP_om.csv",
          row.names = FALSE)
```


# TP v SJ om
```{r}
# Load main results
data <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/ITS/ITS_maaslin/MaaslinResults_cleaned_allSheets_OTUID.xlsx", 
                   sheet = "TP v SJ OM") %>% 
  mutate(OTUID = trimws(as.character(OTUID)))

# Load discriminant OTUs
tpom_discrim <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/ITS/ITS_maaslin/Maaslin_ITS_discrimcomparisons_040725.xlsx", 
                           sheet = "TPOM_discrim") %>% 
  pull(OTUID) %>% trimws()

sjom_discrim <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/ITS/ITS_maaslin/Maaslin_ITS_discrimcomparisons_040725.xlsx", 
                           sheet = "SJOM_discrim") %>% 
  pull(OTUID) %>% trimws()

# Process data
data <- data %>%
  mutate(
    log_q = -log10(qval),
    Enrichment = case_when(
      qval >= 0.05 ~ "Non-Significant",
      coef > 0 ~ "TP Enriched",
      coef < 0 ~ "SJ Enriched"
    ),
    Discriminant = OTUID %in% c(tpom_discrim, sjom_discrim),
    tax_parts = str_split(taxonomy, ";\\s*"),
    family = sapply(tax_parts, function(x) if (length(x) > 0) x[grepl("^f__", x)][1] else NA), # Extract family names
    LabelText = family, # Use only family names for labels
    Label = ifelse(log_q > 7, LabelText, NA) # Apply threshold for labeling
  ) %>%
  filter(!is.na(log_q) & !is.na(coef))  # Remove rows with NA in key columns

# Plot
plot_tp_sj_om <- ggplot(data, aes(x = coef, y = log_q)) +
  geom_point(aes(fill = Enrichment),
             shape = 21, size = 3, stroke = 0, color = "black", alpha = 0.7) +
  geom_point(data = filter(data, Discriminant),
             aes(fill = Enrichment),
             shape = 21, size = 3, stroke = 0.6, color = "black", alpha = 0.7) +
  geom_text_repel(data = filter(data, !is.na(Label)),
                  aes(label = Label),
                  size = 2.5, max.overlaps = 25, segment.color = "gray50") +
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", color = "gray50") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  scale_fill_manual(values = c(
    "TP Enriched" = "#94B594",
    "SJ Enriched" = "#E09351",
    "Non-Significant" = "gray80"
  )) +
  labs(x = "Coefficient", y = "-log10(q-value)", fill = "Enrichment") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    axis.line = element_line(color = "black"),
    axis.ticks = element_line(color = "black"),
    axis.text = element_text(size = 12, color = "black"),
    axis.title = element_text(size = 14, color = "black")
  )

print(plot_tp_sj_om)

# Save plot
ggsave(filename = "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/ITS/ITS_maaslin/ITS_plots/volcano_TP_v_SJ_om.svg",
       plot = plot_tp_sj_om, width = 6, height = 6, dpi = 300, device = "svg")

# Save outliers
outliers_df <- data %>%
  filter(log_q > 7) %>%
  select(OTUID, taxonomy, LabelText, coef, qval, log_q, Enrichment)

write.csv(outliers_df,
          file = "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/ITS/ITS_maaslin/ITS_plots/outliers_TP_v_SJ_om.csv",
          row.names = FALSE)

```


```{r}
install.packages("patchwork")  # If not already installed
library(patchwork)
```

```{r}
# Combine all plots into a 3x3 panel
combined_plot <- (plot_sf_tp_shallow + plot_tp_sj_shallow + plot_sf_sj_shallow) /
                 (plot_sf_tp_deep + plot_sj_tp_deep + plot_sf_sj_deep) /
                 (plot_tp_sf_om + plot_tp_sj_om + plot_sj_sf_om)

# Add labels for rows and columns
combined_plot <- combined_plot +
  plot_annotation(
    title = "Volcano Plots for Site Comparisons",
    subtitle = "Columns: Site Comparisons | Rows: Depths",
    tag_levels = "A"
  ) +
  plot_layout(guides = "collect")  # Combine legends

# Print the combined plot
print(combined_plot)
```

```{r}
# Combine all plots into a 3x3 panel
combined_plot <- (plot_sf_tp_shallow + plot_tp_sj_shallow + plot_sf_sj_shallow) /
                 (plot_sf_tp_deep + plot_sj_tp_deep + plot_sf_sj_deep) /
                 (plot_tp_sf_om + plot_tp_sj_om + plot_sj_sf_om)

# Add labels for rows and columns
combined_plot <- combined_plot +
  plot_annotation(
    title = "Volcano Plots for Site Comparisons",
    subtitle = "Columns: Site Comparisons | Rows: Depths",
    tag_levels = "A"
  ) +
  plot_layout(guides = "collect")  # Combine legends

# Print the combined plot
print(combined_plot)
```


```{r}
print(plot_sf_tp_shallow)
print(plot_tp_sj_shallow)
print(plot_sf_sj_shallow)
print(plot_sf_tp_deep)
print(plot_sj_tp_deep)
print(plot_sf_sj_deep)
print(plot_tp_sf_om)
print(plot_tp_sj_om)
print(plot_sj_sf_om)
```

```{r}
test_combined <- plot_sf_tp_shallow + plot_tp_sj_shallow
print(test_combined)
```


```{r}
# Recreate the plot
plot_sf_tp_om <- ggplot(data, aes(x = coef, y = log_q)) +
  geom_point(aes(fill = Enrichment),
             shape = 21, size = 3, stroke = 0, color = "black", alpha = 0.7) +
  geom_point(data = filter(data, Discriminant),
             aes(fill = Enrichment),
             shape = 21, size = 3, stroke = 0.6, color = "black", alpha = 0.7) +
  geom_text_repel(data = filter(data, !is.na(Label)),
                  aes(label = Label),
                  size = 2.5, max.overlaps = 25, segment.color = "gray50") +
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", color = "gray50") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  scale_fill_manual(values = c(
    "SF Enriched" = "#B75347",
    "TP Enriched" = "#94B594",
    "Non-Significant" = "gray80"
  )) +
  labs(x = "Coefficient", y = "-log10(q-value)", fill = "Enrichment") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    axis.line = element_line(color = "black"),
    axis.ticks = element_line(color = "black"),
    axis.text = element_text(size = 12, color = "black"),
    axis.title = element_text(size = 14, color = "black")
  )

print(plot_sf_tp_om)
```




```{r}
print(plot_sf_tp_om)
```