# ASCC 2023
## Maaslin2 volcano plots redo -> 16S
### 03/05/25
#### Note: Be sure to check the for the discriminant taxa between sites!! Refer to the script located at /Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Rscripts/MaAsLin2/CheckingDiscriminantTaxaByOTU.rmd for additional analysis steps.

**See /Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin2_volcano_plots_wNOTES.md for additional information on analysis steps.**

**Note: This script is for the volcano plots for the Maaslin2 results for the ASCC 2023 16S data.**



**Making volcano plor for SF vs TP shallow**
***I want to overlay the discriminant taxa for all sites within the 0-5cm depth, and I want to do the same with the core taxa across sites & depths***

# Load libraries
```{r}
library(ggplot2)
library(ggrepel)
library(dplyr)
library(readxl)
library(stringr)
```

```{r}
# Read in main dataset
data <- read_excel("/Users/kyasparks/Desktop/MaaslinResults_includingNonSignif_16S_corrected.xlsx", sheet = "SF v TP shallow") %>% as.data.frame()

# Read SF5 discriminant OTUs
discriminant_otus_SF5 <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrim_results2.xlsx", sheet = "SF5_discrim") %>%
  select(OTUID) %>%
  pull()

# Read TP5 discriminant OTUs
discriminant_otus_TP5 <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrim_results2.xlsx", sheet = "TP5_discrim") %>%
  select(OTUID) %>%
  pull()

# Combine both discriminant OTU lists
discriminant_otus <- unique(c(discriminant_otus_SF5, discriminant_otus_TP5))

# Read in core taxa from d70 (core across all sites and depths)
core_taxa_data <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_CoreTaxaAnalysis/CoreTaxa_70_16S_ASCC.xlsx", sheet = "d70_shallow") %>%
  select(OTUID)  # Keep taxonomy info)  # Keep taxonomy info
```


```{r}
# Extract OTUIDs of core taxa
core_taxa <- core_taxa_data$OTUID

# Ensure matching format
data$OTUID <- trimws(as.character(data$OTUID))
discriminant_otus <- trimws(as.character(discriminant_otus))
core_taxa <- trimws(as.character(core_taxa))
```



```{r}
# Convert discriminant OTUs into a dataframe
discriminant_otus_df <- data.frame(OTUID = discriminant_otus)

# Merge with data correctly
data <- data %>%
  left_join(discriminant_otus_df, by = "OTUID") %>%  # Join properly formatted discriminant OTUs
  left_join(core_taxa_data, by = "OTUID") %>%  # Add taxonomy info
  mutate(
    log_q = -log10(qval),  # Standard -log10(qval) transformation
    Enrichment = case_when(
      qval >= 0.05 ~ "Non-Significant",
      coef > 0 ~ "SF Enriched",
      coef < 0 ~ "TP Enriched"
    ),
    
    # ✅ Only highlight if the OTU is discriminant AND significant (qval < 0.05)
    Highlight = ifelse(qval < 0.05 & OTUID %in% discriminant_otus, 
                       "Discriminant Across All Sites in 0-5 cm", "Regular"),
    
    # Core taxa logic
    Core_Taxa = ifelse(OTUID %in% core_taxa, "Core", NA)  
  ) %>%
  mutate(
    # Ensure factors are correctly ordered
    Highlight = factor(Highlight, levels = c("Discriminant Across All Sites in 0-5 cm", "Regular")),
    Core_Taxa = factor(Core_Taxa, levels = c("Core", NA))
  )
```


```{r}
# Identify OTUs that are significant but NOT in discriminant_otus
non_highlighted_otus <- data %>%
  filter(qval < 0.05 & !(OTUID %in% discriminant_otus))

# Base volcano plot
plot <- ggplot(data, aes(x = coef, y = log_q)) +
  
  # Main points with original enrichment color
  geom_point(aes(color = Enrichment, alpha = Enrichment), size = 3) +
  
  # Custom color scheme
  scale_color_manual(values = c(
    "SF Enriched" = "#A1A481",
    "TP Enriched" = "#D97669",
    "Non-Significant" = "#EEC7A0"
  )) +

  # Adjust alpha for non-significant points
  scale_alpha_manual(values = c("Non-Significant" = 1, "SF Enriched" = 1, "TP Enriched" = 1)) +
  
  # 🔥 Overlay discriminant taxa (maintain enrichment color, add black border)
  geom_point(data = data %>% filter(Highlight == "Discriminant Across All Sites in 0-5 cm"),
             aes(x = coef, y = log_q, fill = Enrichment),  
             shape = 21, size = 3, stroke = 0.4, color = "black") +  

  # 🔥 Overlay **significant but non-discriminant OTUs** (e.g., red outline)
  geom_point(data = non_highlighted_otus, 
             aes(x = coef, y = log_q, fill = Enrichment),
             shape = 21, size = 3, stroke = 0.6, color = "red") +  

  # Overlay core taxa (circled points)
  geom_point(data = data %>% filter(!is.na(Core_Taxa)),
             aes(x = coef, y = log_q), shape = 21, fill = NA, size = 3, stroke = 1, color = "green") +

  # geom_text_repel(data = data %>% filter(!is.na(Core_Taxa)), 
  #                 aes(x = coef, y = log_q, label = OTUID), 
  #                 size = 3, box.padding = 0.5, segment.color = "gray") +
  
  
  # Apply same colors to `fill`
  scale_fill_manual(values = c(
    "SF Enriched" = "#A1A481",
    "TP Enriched" = "#D97669",
    "Non-Significant" = "#EEC7A0"
  )) +
  
  # Threshold lines
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", color = "gray") +

  # Labels & theme
  labs(x = "Coefficient", 
       y = "-log10(q-value)", 
       color = "Enrichment", 
       fill = "Enrichment", 
       shape = "Highlighted Taxa") +  

  # Ensure shape legend only applies to discriminant taxa
  scale_shape_manual(values = c("Discriminant Across All Sites in 0-5 cm" = 21)) +  

  theme_minimal() +
  theme(legend.position = "top")

# Show plot
print(plot)
```

```{r}
# Save the volcano plot as a square SVG file
ggsave(filename = "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Plots/volcano_plot_SFTP5.svg",
       plot = plot, 
       width = 6, height = 6, dpi = 300, device = "svg")
```

```{r}
# Print OTUIDs of Core Taxa (Circled)
core_otus <- data %>%
  filter(Core_Taxa == "Core") %>%
  select(OTUID)

# View the OTUIDs that were circled
print(core_otus)

# Count how many OTUs were circled
cat("Number of Core Taxa (Circled):", nrow(core_otus), "\n")

# Save the OTUIDs to a CSV file (optional)
write.csv(core_otus, "core_OTUIDssftp5.csv", row.names = FALSE)

# Save to CSV (optional)
write.csv(non_highlighted_otus, "non_highlighted_OTUIDs_sftp5.csv", row.names = FALSE)
```

**Making volcano plot for SF vs SJ shallow**


```{r}
# Read in main dataset
data <- read_excel("/Users/kyasparks/Desktop/MaaslinResults_includingNonSignif_16S_corrected.xlsx", sheet = "SF v SJ shallow") %>% as.data.frame()

# Read SF5 discriminant OTUs
discriminant_otus_SF5 <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrim_results2.xlsx", sheet = "SF5_discrim") %>%
  select(OTUID) %>%
  pull()

# Read TP5 discriminant OTUs
discriminant_otus_SJ5 <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrim_results2.xlsx", sheet = "SJ5_discrim") %>%
  select(OTUID) %>%
  pull()

# Combine both discriminant OTU lists
discriminant_otus <- unique(c(discriminant_otus_SF5, discriminant_otus_SJ5))

# Read in core taxa from d70 (core across all sites and depths)
core_taxa_data <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_CoreTaxaAnalysis/CoreTaxa_70_16S_ASCC.xlsx", sheet = "d70_shallow") %>%
  select(OTUID)  # Keep taxonomy info)  # Keep taxonomy info
```


```{r}
# Extract OTUIDs of core taxa
core_taxa <- core_taxa_data$OTUID

# Ensure matching format
data$OTUID <- trimws(as.character(data$OTUID))
discriminant_otus <- trimws(as.character(discriminant_otus))
core_taxa <- trimws(as.character(core_taxa))
```

```{r}
# Convert discriminant OTUs into a dataframe
discriminant_otus_df <- data.frame(OTUID = discriminant_otus)

# Merge with data correctly
data <- data %>%
  left_join(discriminant_otus_df, by = "OTUID") %>%  # Join properly formatted discriminant OTUs
  left_join(core_taxa_data, by = "OTUID") %>%  # Add taxonomy info
  mutate(
    log_q = -log10(qval),  # Standard -log10(qval) transformation
    Enrichment = case_when(
      qval >= 0.05 ~ "Non-Significant",
      coef > 0 ~ "SF Enriched",
      coef < 0 ~ "SJ Enriched"
    ),
    
    # ✅ Only highlight if the OTU is discriminant AND significant (qval < 0.05)
    Highlight = ifelse(qval < 0.05 & OTUID %in% discriminant_otus, 
                       "Discriminant Across All Sites in 0-5 cm", "Regular"),
    
    # Core taxa logic
    Core_Taxa = ifelse(OTUID %in% core_taxa, "Core", NA)  
  ) %>%
  mutate(
    # Ensure factors are correctly ordered
    Highlight = factor(Highlight, levels = c("Discriminant Across All Sites in 0-5 cm", "Regular")),
    Core_Taxa = factor(Core_Taxa, levels = c("Core", NA))
  )
```

```{r}
# Identify OTUs that are significant but NOT in discriminant_otus
non_highlighted_otus <- data %>%
  filter(qval < 0.05 & !(OTUID %in% discriminant_otus))

# Base volcano plot
plot <- ggplot(data, aes(x = coef, y = log_q)) +
  
  # Main points with original enrichment color
  geom_point(aes(color = Enrichment, alpha = Enrichment), size = 3) +
  
  # Custom color scheme
  scale_color_manual(values = c(
    "SF Enriched" = "#A1A481",
    "SJ Enriched" = "#D97669",
    "Non-Significant" = "#EEC7A0"
  )) +#BD7C96

  # Adjust alpha for non-significant points
  scale_alpha_manual(values = c("Non-Significant" = 1, "SF Enriched" = 1, "SJ Enriched" = 1)) +
  
  # 🔥 Overlay discriminant taxa (maintain enrichment color, add black border)
  geom_point(data = data %>% filter(Highlight == "Discriminant Across All Sites in 0-5 cm"),
             aes(x = coef, y = log_q, fill = Enrichment),  
             shape = 21, size = 3, stroke = 0.4, color = "black") +  

  # Overlay **significant but non-discriminant OTUs** (e.g., red outline)
  geom_point(data = non_highlighted_otus, 
             aes(x = coef, y = log_q, fill = Enrichment),
             shape = 21, size = 3, stroke = 0.6, color = "red") +  

  # Overlay core taxa (circled points)
  geom_point(data = data %>% filter(!is.na(Core_Taxa)),
             aes(x = coef, y = log_q), shape = 21, fill = NA, size = 3, stroke = 1, color = "green") +

  # geom_text_repel(data = data %>% filter(!is.na(Core_Taxa)), 
  #                 aes(x = coef, y = log_q, label = OTUID), 
  #                 size = 3, box.padding = 0.5, segment.color = "gray") +
  
  
  # Apply same colors to `fill`
  scale_fill_manual(values = c(
    "SF Enriched" = "#A1A481",
    "SJ Enriched" = "#BD7C96",
    "Non-Significant" = "#EEC7A0"
  )) +
  
  # Threshold lines
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", color = "gray") +

  # Labels & theme
  labs(x = "Coefficient", 
       y = "-log10(q-value)", 
       color = "Enrichment", 
       fill = "Enrichment", 
       shape = "Highlighted Taxa") +  

  # Ensure shape legend only applies to discriminant taxa
  scale_shape_manual(values = c("Discriminant Across All Sites in 0-5 cm" = 21)) +  

  theme_minimal() +
  theme(legend.position = "top")

# Show plot
print(plot)
```


```{r}
# Save the volcano plot as a square SVG file
ggsave(filename = "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Plots/volcano_plot_SFSJ5.svg",
       plot = plot, 
       width = 6, height = 6, dpi = 300, device = "svg")
```

```{r}
# Print OTUIDs of Core Taxa (Circled)
core_otus <- data %>%
  filter(Core_Taxa == "Core") %>%
  select(OTUID)

# View the OTUIDs that were circled
print(core_otus)

# Count how many OTUs were circled
cat("Number of Core Taxa (Circled):", nrow(core_otus), "\n")

# Save the OTUIDs to a CSV file (optional)
write.csv(core_otus, "core_OTUIDs_SFSJ.csv", row.names = FALSE)

print(non_highlighted_otus)

# Save to CSV (optional)
write.csv(non_highlighted_otus, "non_highlighted_OTUIDs_sfsj5.csv", row.names = FALSE)

# highlighted_otus <- data %>% 
#   filter(Highlight == "Discriminant Across All Sites in 0-5 cm") %>% 
#   select(OTUID)

#print(highlighted_otus)

#write.csv(highlighted_otus, "highlighted_otus_SFSJ.csv", row.names = FALSE)
```


**Making volcano plot for TP vs SJ shallow**


```{r}
# Read in main dataset
data <- read_excel("/Users/kyasparks/Desktop/MaaslinResults_includingNonSignif_16S_corrected.xlsx", sheet = "TP v SJ shallow") %>% as.data.frame()

# Read SF5 discriminant OTUs
discriminant_otus_TP5 <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrim_results2.xlsx", sheet = "TP5_discrim") %>%
  select(OTUID) %>%
  pull()

# Read TP5 discriminant OTUs
discriminant_otus_SJ5 <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrim_results2.xlsx", sheet = "SJ5_discrim") %>%
  select(OTUID) %>%
  pull()

# Combine both discriminant OTU lists
discriminant_otus <- unique(c(discriminant_otus_TP5, discriminant_otus_SJ5))

# Read in core taxa from d70 (core across all sites and depths)
core_taxa_data <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_CoreTaxaAnalysis/CoreTaxa_70_16S_ASCC.xlsx", sheet = "d70_shallow") %>%
  select(OTUID)  # Keep taxonomy info)  # Keep taxonomy info
```


```{r}
# Extract OTUIDs of core taxa
core_taxa <- core_taxa_data$OTUID

# Ensure matching format
data$OTUID <- trimws(as.character(data$OTUID))
discriminant_otus <- trimws(as.character(discriminant_otus))
core_taxa <- trimws(as.character(core_taxa))
```


```{r}
# Convert discriminant OTUs into a dataframe
discriminant_otus_df <- data.frame(OTUID = discriminant_otus)

# Merge with data correctly
data <- data %>%
  left_join(discriminant_otus_df, by = "OTUID") %>%  # Join properly formatted discriminant OTUs
  left_join(core_taxa_data, by = "OTUID") %>%  # Add taxonomy info
  mutate(
    log_q = -log10(qval),  # Standard -log10(qval) transformation
    Enrichment = case_when(
      qval >= 0.05 ~ "Non-Significant",
      coef > 0 ~ "TP Enriched",
      coef < 0 ~ "SJ Enriched"
    ),
    
    # ✅ Only highlight if the OTU is discriminant AND significant (qval < 0.05)
    Highlight = ifelse(qval < 0.05 & OTUID %in% discriminant_otus, 
                       "Discriminant Across All Sites in 0-5 cm", "Regular"),
    
    # Core taxa logic
    Core_Taxa = ifelse(OTUID %in% core_taxa, "Core", NA)  
  ) %>%
  mutate(
    # Ensure factors are correctly ordered
    Highlight = factor(Highlight, levels = c("Discriminant Across All Sites in 0-5 cm", "Regular")),
    Core_Taxa = factor(Core_Taxa, levels = c("Core", NA))
  )
```

```{r}
# Identify OTUs that are significant but NOT in discriminant_otus
non_highlighted_otus <- data %>%
  filter(qval < 0.05 & !(OTUID %in% discriminant_otus))

# Base volcano plot
plot <- ggplot(data, aes(x = coef, y = log_q)) +
  
  # Main points with original enrichment color
  geom_point(aes(color = Enrichment, alpha = Enrichment), size = 3) +
  
  # Custom color scheme
  scale_color_manual(values = c(
    "TP Enriched" = "#A1A481",
    "SJ Enriched" = "#BD7C96",
    "Non-Significant" = "#EEC7A0"
  )) +

  # Adjust alpha for non-significant points
  scale_alpha_manual(values = c("Non-Significant" = 1, "TP Enriched" = 1, "SJ Enriched" = 1)) +
  
  # 🔥 Overlay discriminant taxa (maintain enrichment color, add black border)
  geom_point(data = data %>% filter(Highlight == "Discriminant Across All Sites in 0-5 cm"),
             aes(x = coef, y = log_q, fill = Enrichment),  
             shape = 21, size = 3, stroke = 0.4, color = "black") +  

  # Overlay **significant but non-discriminant OTUs** (e.g., red outline)
  geom_point(data = non_highlighted_otus, 
             aes(x = coef, y = log_q, fill = Enrichment),
             shape = 21, size = 3, stroke = 0.6, color = "red") +  

  # Overlay core taxa (circled points)
  geom_point(data = data %>% filter(!is.na(Core_Taxa)),
             aes(x = coef, y = log_q), shape = 21, fill = NA, size = 3, stroke = 1, color = "green") +

  # geom_text_repel(data = data %>% filter(!is.na(Core_Taxa)), 
  #                 aes(x = coef, y = log_q, label = OTUID), 
  #                 size = 3, box.padding = 0.5, segment.color = "gray") +
  
  
  # Apply same colors to `fill`
  scale_fill_manual(values = c(
    "TP Enriched" = "#A1A481",
    "SJ Enriched" = "#BD7C96",
    "Non-Significant" = "#EEC7A0"
  )) +
  
  # Threshold lines
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", color = "gray") +

  # Labels & theme
  labs(x = "Coefficient", 
       y = "-log10(q-value)", 
       color = "Enrichment", 
       fill = "Enrichment", 
       shape = "Highlighted Taxa") +  

  # Ensure shape legend only applies to discriminant taxa
  scale_shape_manual(values = c("Discriminant Across All Sites in 0-5 cm" = 21)) +  

  theme_minimal() +
  theme(legend.position = "top")

# Show plot
print(plot)
```

```{r}
# Save the volcano plot as a square SVG file
ggsave(filename = "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Plots/volcano_plot_TPSJ5.svg",
       plot = plot, 
       width = 6, height = 6, dpi = 300, device = "svg")
```

```{r}
# Print OTUIDs of Core Taxa (Circled)
core_otus <- data %>%
  filter(Core_Taxa == "Core") %>%
  select(OTUID)

# View the OTUIDs that were circled
print(core_otus)

# Count how many OTUs were circled
cat("Number of Core Taxa (Circled):", nrow(core_otus), "\n")

# Save the OTUIDs to a CSV file (optional)
write.csv(core_otus, "core_OTUIDs_tpsj5.csv", row.names = FALSE)

print(non_highlighted_otus)

# Save to CSV (optional)
write.csv(non_highlighted_otus, "non_highlighted_OTUIDs_tpsj5.csv", row.names = FALSE)

# highlighted_otus <- data %>% 
#   filter(Highlight == "Discriminant Across All Sites in 0-5 cm") %>% 
#   select(OTUID)

# print(highlighted_otus)

# write.csv(highlighted_otus, "highlighted_otus_tpsj5.csv", row.names = FALSE)
```















**SF v TP deep**

```{r}
# Read in main dataset
data <- read_excel("/Users/kyasparks/Desktop/MaaslinResults_includingNonSignif_16S_corrected.xlsx", sheet = "SF v TP deep") %>% as.data.frame()

# Read SF5 discriminant OTUs
discriminant_otus_SF15 <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrim_results2.xlsx", sheet = "SF15_discrim") %>%
  select(OTUID) %>%
  pull()

# Read TP5 discriminant OTUs
discriminant_otus_TP15 <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrim_results2.xlsx", sheet = "TP15_discrim") %>%
  select(OTUID) %>%
  pull()

# Combine both discriminant OTU lists
discriminant_otus <- unique(c(discriminant_otus_SF15, discriminant_otus_TP15))

# Read in core taxa from d70 (core across all sites and depths)
core_taxa_data <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_CoreTaxaAnalysis/CoreTaxa_70_16S_ASCC.xlsx", sheet = "d70_deep") %>%
  select(OTUID)  # Keep taxonomy info  # Keep taxonomy info
```


```{r}
# Extract OTUIDs of core taxa
core_taxa <- core_taxa_data$OTUID

# Ensure matching format
data$OTUID <- trimws(as.character(data$OTUID))
discriminant_otus <- trimws(as.character(discriminant_otus))
core_taxa <- trimws(as.character(core_taxa))
```



```{r}
# Convert discriminant OTUs into a dataframe
discriminant_otus_df <- data.frame(OTUID = discriminant_otus)

# Merge with data correctly
data <- data %>%
  left_join(discriminant_otus_df, by = "OTUID") %>%  # Join properly formatted discriminant OTUs
  left_join(core_taxa_data, by = "OTUID") %>%  # Add taxonomy info
  mutate(
    log_q = -log10(qval),  # Standard -log10(qval) transformation
    Enrichment = case_when(
      qval >= 0.05 ~ "Non-Significant",
      coef > 0 ~ "SF Enriched",
      coef < 0 ~ "TP Enriched"
    ),
    
    # ✅ Only highlight if the OTU is discriminant AND significant (qval < 0.05)
    Highlight = ifelse(qval < 0.05 & OTUID %in% discriminant_otus, 
                       "Discriminant Across All Sites in 5-15 cm", "Regular"),
    
    # Core taxa logic
    Core_Taxa = ifelse(OTUID %in% core_taxa, "Core", NA)  
  ) %>%
  mutate(
    # Ensure factors are correctly ordered
    Highlight = factor(Highlight, levels = c("Discriminant Across All Sites in 5-15 cm", "Regular")),
    Core_Taxa = factor(Core_Taxa, levels = c("Core", NA))
  )
```


```{r}
# Identify OTUs that are significant but NOT in discriminant_otus
non_highlighted_otus <- data %>%
  filter(qval < 0.05 & !(OTUID %in% discriminant_otus))

# Base volcano plot
plot <- ggplot(data, aes(x = coef, y = log_q)) +
  
  # Main points with original enrichment color
  geom_point(aes(color = Enrichment, alpha = Enrichment), size = 3) +
  
  # Custom color scheme
  scale_color_manual(values = c(
    "SF Enriched" = "#A1A481",
    "TP Enriched" = "#D97669",
    "Non-Significant" = "#EEC7A0"
  )) +

  # Adjust alpha for non-significant points
  scale_alpha_manual(values = c("Non-Significant" = 1, "SF Enriched" = 1, "TP Enriched" = 1)) +
  
  # 🔥 Overlay discriminant taxa (maintain enrichment color, add black border)
  geom_point(data = data %>% filter(Highlight == "Discriminant Across All Sites in 5-15 cm"),
             aes(x = coef, y = log_q, fill = Enrichment),  
             shape = 21, size = 3, stroke = 0.4, color = "black") +  

  # 🔥 Overlay **significant but non-discriminant OTUs** (e.g., red outline)
  geom_point(data = non_highlighted_otus, 
             aes(x = coef, y = log_q, fill = Enrichment),
             shape = 21, size = 3, stroke = 0.6, color = "red") +  

  # Overlay core taxa (circled points)
  geom_point(data = data %>% filter(!is.na(Core_Taxa)),
             aes(x = coef, y = log_q), shape = 21, fill = NA, size = 3, stroke = 1, color = "green") +

  geom_text_repel(data = data %>% filter(!is.na(Core_Taxa)), 
                  aes(x = coef, y = log_q, label = OTUID), 
                  size = 3, box.padding = 0.5, segment.color = "gray") +
  
  
  # Apply same colors to `fill`
  scale_fill_manual(values = c(
    "SF Enriched" = "#A1A481",
    "TP Enriched" = "#D97669",
    "Non-Significant" = "#EEC7A0"
  )) +
  
  # Threshold lines
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", color = "gray") +

  # Labels & theme
  labs(x = "Coefficient", 
       y = "-log10(q-value)", 
       color = "Enrichment", 
       fill = "Enrichment", 
       shape = "Highlighted Taxa") +  

  # Ensure shape legend only applies to discriminant taxa
  scale_shape_manual(values = c("Discriminant Across All Sites in 5-15 cm" = 21)) +  

  theme_minimal() +
  theme(legend.position = "top")

# Show plot
print(plot)
```

```{r}
# Save the volcano plot as a square SVG file
ggsave(filename = "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Plots/volcano_plot_SFTP15_coredeep.svg",
       plot = plot, 
       width = 6, height = 6, dpi = 300, device = "svg")
```

```{r}
# Print OTUIDs of Core Taxa (Circled)
core_otus <- data %>%
  filter(Core_Taxa == "Core") %>%
  select(OTUID)

# View the OTUIDs that were circled
print(core_otus)

# Count how many OTUs were circled
cat("Number of Core Taxa (Circled):", nrow(core_otus), "\n")

# Save the OTUIDs to a CSV file (optional)
write.csv(core_otus, "core_OTUIDs_sftp15.csv", row.names = FALSE)

print(non_highlighted_otus)

# Save to CSV (optional)
write.csv(non_highlighted_otus, "non_highlighted_OTUIDs_sftp15.csv", row.names = FALSE)
```








**Making volcano plot for SF vs SJ deep**


```{r}
# Read in main dataset
data <- read_excel("/Users/kyasparks/Desktop/MaaslinResults_includingNonSignif_16S_corrected.xlsx", sheet = "SJ v SF deep") %>% as.data.frame()

# Read SF5 discriminant OTUs
discriminant_otus_SF15 <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrim_results2.xlsx", sheet = "SF15_discrim") %>%
  select(OTUID) %>%
  pull()

# Read TP5 discriminant OTUs
discriminant_otus_SJ15 <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrim_results2.xlsx", sheet = "SJ15_discrim") %>%
  select(OTUID) %>%
  pull()

# Combine both discriminant OTU lists
discriminant_otus <- unique(c(discriminant_otus_SF15, discriminant_otus_SJ15))

# Read in core taxa from d70 (core across all sites and depths)
core_taxa_data <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_CoreTaxaAnalysis/CoreTaxa_70_16S_ASCC.xlsx", sheet = "d70_deep") %>%
  select(OTUID)  # Keep taxonomy info
```


```{r}
# Extract OTUIDs of core taxa
core_taxa <- core_taxa_data$OTUID

# Ensure matching format
data$OTUID <- trimws(as.character(data$OTUID))
discriminant_otus <- trimws(as.character(discriminant_otus))
core_taxa <- trimws(as.character(core_taxa))
```

```{r}
# Convert discriminant OTUs into a dataframe
discriminant_otus_df <- data.frame(OTUID = discriminant_otus)

# Merge with data correctly
data <- data %>%
  left_join(discriminant_otus_df, by = "OTUID") %>%  # Join properly formatted discriminant OTUs
  left_join(core_taxa_data, by = "OTUID") %>%  # Add taxonomy info
  mutate(
    log_q = -log10(qval),  # Standard -log10(qval) transformation
    Enrichment = case_when(
      qval >= 0.05 ~ "Non-Significant",
      coef > 0 ~ "SF Enriched",
      coef < 0 ~ "SJ Enriched"
    ),
    
    # ✅ Only highlight if the OTU is discriminant AND significant (qval < 0.05)
    Highlight = ifelse(qval < 0.05 & OTUID %in% discriminant_otus, 
                       "Discriminant Across All Sites in 5-15 cm", "Regular"),
    
    # Core taxa logic
    Core_Taxa = ifelse(OTUID %in% core_taxa, "Core", NA)  
  ) %>%
  mutate(
    # Ensure factors are correctly ordered
    Highlight = factor(Highlight, levels = c("Discriminant Across All Sites in 5-15 cm", "Regular")),
    Core_Taxa = factor(Core_Taxa, levels = c("Core", NA))
  )
```

```{r}
# Identify OTUs that are significant but NOT in discriminant_otus
non_highlighted_otus <- data %>%
  filter(qval < 0.05 & !(OTUID %in% discriminant_otus))

# Base volcano plot
plot <- ggplot(data, aes(x = coef, y = log_q)) +
  
  # Main points with original enrichment color
  geom_point(aes(color = Enrichment, alpha = Enrichment), size = 3) +
  
  # Custom color scheme
  scale_color_manual(values = c(
    "SF Enriched" = "#A1A481",
    "SJ Enriched" = "#BD7C96",
    "Non-Significant" = "#EEC7A0"
  )) +

  # Adjust alpha for non-significant points
  scale_alpha_manual(values = c("Non-Significant" = 1, "SF Enriched" = 1, "SJ Enriched" = 1)) +
  
  # 🔥 Overlay discriminant taxa (maintain enrichment color, add black border)
  geom_point(data = data %>% filter(Highlight == "Discriminant Across All Sites in 5-15 cm"),
             aes(x = coef, y = log_q, fill = Enrichment),  
             shape = 21, size = 3, stroke = 0.4, color = "black") +  

  # Overlay **significant but non-discriminant OTUs** (e.g., red outline)
  geom_point(data = non_highlighted_otus, 
             aes(x = coef, y = log_q, fill = Enrichment),
             shape = 21, size = 3, stroke = 0.6, color = "red") +  

  # Overlay core taxa (circled points)
  geom_point(data = data %>% filter(!is.na(Core_Taxa)),
             aes(x = coef, y = log_q), shape = 21, fill = NA, size = 3, stroke = 1, color = "green") +

  geom_text_repel(data = data %>% filter(!is.na(Core_Taxa)), 
                  aes(x = coef, y = log_q, label = OTUID), 
                  size = 3, box.padding = 0.5, segment.color = "gray") +
  
  
  # Apply same colors to `fill`
  scale_fill_manual(values = c(
    "SF Enriched" = "#A1A481",
    "SJ Enriched" = "#BD7C96",
    "Non-Significant" = "#EEC7A0"
  )) +
  
  # Threshold lines
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", color = "gray") +

  # Labels & theme
  labs(x = "Coefficient", 
       y = "-log10(q-value)", 
       color = "Enrichment", 
       fill = "Enrichment", 
       shape = "Highlighted Taxa") +  

  # Ensure shape legend only applies to discriminant taxa
  scale_shape_manual(values = c("Discriminant Across All Sites in 5-15 cm" = 21)) +  

  theme_minimal() +
  theme(legend.position = "top")

# Show plot
print(plot)
```


```{r}
# Save the volcano plot as a square SVG file
ggsave(filename = "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Plots/volcano_plot_SFSJ15_core15.svg",
       plot = plot, 
       width = 6, height = 6, dpi = 300, device = "svg")
```

```{r}
# Print OTUIDs of Core Taxa (Circled)
core_otus <- data %>%
  filter(Core_Taxa == "Core") %>%
  select(OTUID)

# View the OTUIDs that were circled
print(core_otus)

# Count how many OTUs were circled
cat("Number of Core Taxa (Circled):", nrow(core_otus), "\n")

# Save the OTUIDs to a CSV file (optional)
write.csv(core_otus, "core_OTUIDs_SFSJ15.csv", row.names = FALSE)

print(non_highlighted_otus)

# Save to CSV (optional)
write.csv(non_highlighted_otus, "non_highlighted_OTUIDs_sfsj15.csv", row.names = FALSE)
```


**Making volcano plot for TP vs SJ deep**


```{r}
# Read in main dataset
data <- read_excel("/Users/kyasparks/Desktop/MaaslinResults_includingNonSignif_16S_corrected.xlsx", sheet = "TP v SJ deep") %>% as.data.frame()

# Read SF5 discriminant OTUs
discriminant_otus_TP15 <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrim_results2.xlsx", sheet = "TP15_discrim") %>%
  select(OTUID) %>%
  pull()

# Read TP5 discriminant OTUs
discriminant_otus_SJ15 <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrim_results2.xlsx", sheet = "SJ15_discrim") %>%
  select(OTUID) %>%
  pull()

# Combine both discriminant OTU lists
discriminant_otus <- unique(c(discriminant_otus_TP15, discriminant_otus_SJ15))

# Read in core taxa from d70 (core across all sites and depths)
core_taxa_data <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_CoreTaxaAnalysis/CoreTaxa_70_16S_ASCC.xlsx", sheet = "d70_deep") %>%
  select(OTUID)  # Keep taxonomy info  # Keep taxonomy info
```


```{r}
# Extract OTUIDs of core taxa
core_taxa <- core_taxa_data$OTUID

# Ensure matching format
data$OTUID <- trimws(as.character(data$OTUID))
discriminant_otus <- trimws(as.character(discriminant_otus))
core_taxa <- trimws(as.character(core_taxa))
```


```{r}
# Convert discriminant OTUs into a dataframe
discriminant_otus_df <- data.frame(OTUID = discriminant_otus)

# Merge with data correctly
data <- data %>%
  left_join(discriminant_otus_df, by = "OTUID") %>%  # Join properly formatted discriminant OTUs
  left_join(core_taxa_data, by = "OTUID") %>%  # Add taxonomy info
  mutate(
    log_q = -log10(qval),  # Standard -log10(qval) transformation
    Enrichment = case_when(
      qval >= 0.05 ~ "Non-Significant",
      coef > 0 ~ "TP Enriched",
      coef < 0 ~ "SJ Enriched"
    ),
    
    # ✅ Only highlight if the OTU is discriminant AND significant (qval < 0.05)
    Highlight = ifelse(qval < 0.05 & OTUID %in% discriminant_otus, 
                       "Discriminant Across All Sites in 5-15 cm", "Regular"),
    
    # Core taxa logic
    Core_Taxa = ifelse(OTUID %in% core_taxa, "Core", NA)  
  ) %>%
  mutate(
    # Ensure factors are correctly ordered
    Highlight = factor(Highlight, levels = c("Discriminant Across All Sites in 5-15 cm", "Regular")),
    Core_Taxa = factor(Core_Taxa, levels = c("Core", NA))
  )
```

```{r}
# Identify OTUs that are significant but NOT in discriminant_otus
non_highlighted_otus <- data %>%
  filter(qval < 0.05 & !(OTUID %in% discriminant_otus))

# Base volcano plot
plot <- ggplot(data, aes(x = coef, y = log_q)) +
  
  # Main points with original enrichment color
  geom_point(aes(color = Enrichment, alpha = Enrichment), size = 3) +
  
  # Custom color scheme
  scale_color_manual(values = c(
    "TP Enriched" = "#A1A481",
    "SJ Enriched" = "#BD7C96",
    "Non-Significant" = "#EEC7A0"
  )) +

  # Adjust alpha for non-significant points
  scale_alpha_manual(values = c("Non-Significant" = 1, "TP Enriched" = 1, "SJ Enriched" = 1)) +
  
  # 🔥 Overlay discriminant taxa (maintain enrichment color, add black border)
  geom_point(data = data %>% filter(Highlight == "Discriminant Across All Sites in 5-15 cm"),
             aes(x = coef, y = log_q, fill = Enrichment),  
             shape = 21, size = 3, stroke = 0.4, color = "black") +  

  # Overlay **significant but non-discriminant OTUs** (e.g., red outline)
  geom_point(data = non_highlighted_otus, 
             aes(x = coef, y = log_q, fill = Enrichment),
             shape = 21, size = 3, stroke = 0.6, color = "red") +  

  # Overlay core taxa (circled points)
  geom_point(data = data %>% filter(!is.na(Core_Taxa)),
             aes(x = coef, y = log_q), shape = 21, fill = NA, size = 3, stroke = 1, color = "green") +

  geom_text_repel(data = data %>% filter(!is.na(Core_Taxa)), 
                  aes(x = coef, y = log_q, label = OTUID), 
                  size = 3, box.padding = 0.5, segment.color = "gray") +
  
  
  # Apply same colors to `fill`
  scale_fill_manual(values = c(
    "TP Enriched" = "#A1A481",
    "SJ Enriched" = "#BD7C96",
    "Non-Significant" = "#EEC7A0"
  )) +
  
  # Threshold lines
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", color = "gray") +

  # Labels & theme
  labs(x = "Coefficient", 
       y = "-log10(q-value)", 
       color = "Enrichment", 
       fill = "Enrichment", 
       shape = "Highlighted Taxa") +  

  # Ensure shape legend only applies to discriminant taxa
  scale_shape_manual(values = c("Discriminant Across All Sites in 5-15 cm" = 21)) +  

  theme_minimal() +
  theme(legend.position = "top")

# Show plot
print(plot)
```

```{r}
# Save the volcano plot as a square SVG file
ggsave(filename = "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Plots/volcano_plot_TPSJ15_core15.svg",
       plot = plot, 
       width = 6, height = 6, dpi = 300, device = "svg")
```

```{r}
# Print OTUIDs of Core Taxa (Circled)
core_otus <- data %>%
  filter(Core_Taxa == "Core") %>%
  select(OTUID)

# View the OTUIDs that were circled
print(core_otus)

# Count how many OTUs were circled
cat("Number of Core Taxa (Circled):", nrow(core_otus), "\n")

# Save the OTUIDs to a CSV file (optional)
write.csv(core_otus, "core_OTUIDs_tpsj15.csv", row.names = FALSE)

print(non_highlighted_otus)

# Save to CSV (optional)
write.csv(non_highlighted_otus, "non_highlighted_OTUIDs_tpsj15.csv", row.names = FALSE)

highlighted_otus <- data %>% 
  filter(Highlight == "Discriminant Across All Sites in 0-5 cm") %>% 
  select(OTUID)

print(highlighted_otus)

write.csv(highlighted_otus, "highlighted_otus_tpsj15.csv", row.names = FALSE)
```






**SF v TP OM**

```{r}
# Read in main dataset
data <- read_excel("/Users/kyasparks/Desktop/MaaslinResults_includingNonSignif_16S_corrected.xlsx", sheet = "SF v TP OM") %>% as.data.frame()

# Read SF5 discriminant OTUs
discriminant_otus_SFOM <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrim_results2.xlsx", sheet = "SFOM_discrim") %>%
  select(OTUID) %>%
  pull()

# Read TP5 discriminant OTUs
discriminant_otus_TPOM <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrim_results2.xlsx", sheet = "TPOM_discrim") %>%
  select(OTUID) %>%
  pull()

# Combine both discriminant OTU lists
discriminant_otus <- unique(c(discriminant_otus_SFOM, discriminant_otus_TPOM))

# Read in core taxa from d70 (core across all sites and depths)
core_taxa_data <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_CoreTaxaAnalysis/CoreTaxa_70_16S_ASCC.xlsx", sheet = "d70_OM") %>%
  select(OTUID)  # Keep taxonomy info  # Keep taxonomy info
```


```{r}
# Extract OTUIDs of core taxa
core_taxa <- core_taxa_data$OTUID

# Ensure matching format
data$OTUID <- trimws(as.character(data$OTUID))
discriminant_otus <- trimws(as.character(discriminant_otus))
core_taxa <- trimws(as.character(core_taxa))
```



```{r}
# Convert discriminant OTUs into a dataframe
discriminant_otus_df <- data.frame(OTUID = discriminant_otus)

# Merge with data correctly
data <- data %>%
  left_join(discriminant_otus_df, by = "OTUID") %>%  # Join properly formatted discriminant OTUs
  left_join(core_taxa_data, by = "OTUID") %>%  # Add taxonomy info
  mutate(
    log_q = -log10(qval),  # Standard -log10(qval) transformation
    Enrichment = case_when(
      qval >= 0.05 ~ "Non-Significant",
      coef > 0 ~ "SF Enriched",
      coef < 0 ~ "TP Enriched"
    ),
    
    # ✅ Only highlight if the OTU is discriminant AND significant (qval < 0.05)
    Highlight = ifelse(qval < 0.05 & OTUID %in% discriminant_otus, 
                       "Discriminant Across All Sites in OM", "Regular"),
    
    # Core taxa logic
    Core_Taxa = ifelse(OTUID %in% core_taxa, "Core", NA)  
  ) %>%
  mutate(
    # Ensure factors are correctly ordered
    Highlight = factor(Highlight, levels = c("Discriminant Across All Sites in OM", "Regular")),
    Core_Taxa = factor(Core_Taxa, levels = c("Core", NA))
  )
```


```{r}
# Identify OTUs that are significant but NOT in discriminant_otus
non_highlighted_otus <- data %>%
  filter(qval < 0.05 & !(OTUID %in% discriminant_otus))

# Base volcano plot
plot <- ggplot(data, aes(x = coef, y = log_q)) +
  
  # Main points with original enrichment color
  geom_point(aes(color = Enrichment, alpha = Enrichment), size = 3) +
  
  # Custom color scheme
  scale_color_manual(values = c(
    "SF Enriched" = "#A1A481",
    "TP Enriched" = "#D97669",
    "Non-Significant" = "#EEC7A0"
  )) +

  # Adjust alpha for non-significant points
  scale_alpha_manual(values = c("Non-Significant" = 1, "SF Enriched" = 1, "TP Enriched" = 1)) +
  
  # 🔥 Overlay discriminant taxa (maintain enrichment color, add black border)
  geom_point(data = data %>% filter(Highlight == "Discriminant Across All Sites in OM"),
             aes(x = coef, y = log_q, fill = Enrichment),  
             shape = 21, size = 3, stroke = 0.4, color = "black") +  

  # 🔥 Overlay **significant but non-discriminant OTUs** (e.g., red outline)
  geom_point(data = non_highlighted_otus, 
             aes(x = coef, y = log_q, fill = Enrichment),
             shape = 21, size = 3, stroke = 0.6, color = "red") +  

  # Overlay core taxa (circled points)
  geom_point(data = data %>% filter(!is.na(Core_Taxa)),
             aes(x = coef, y = log_q), shape = 21, fill = NA, size = 3, stroke = 1, color = "green") +

  geom_text_repel(data = data %>% filter(!is.na(Core_Taxa)), 
                  aes(x = coef, y = log_q, label = OTUID), 
                  size = 3, box.padding = 0.5, segment.color = "gray") +
  
  
  # Apply same colors to `fill`
  scale_fill_manual(values = c(
    "SF Enriched" = "#A1A481",
    "TP Enriched" = "#D97669",
    "Non-Significant" = "#EEC7A0"
  )) +
  
  # Threshold lines
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", color = "gray") +

  # Labels & theme
  labs(x = "Coefficient", 
       y = "-log10(q-value)", 
       color = "Enrichment", 
       fill = "Enrichment", 
       shape = "Highlighted Taxa") +  

  # Ensure shape legend only applies to discriminant taxa
  scale_shape_manual(values = c("Discriminant Across All Sites in OM" = 21)) +  

  theme_minimal() +
  theme(legend.position = "top")

# Show plot
print(plot)
```

```{r}
# Save the volcano plot as a square SVG file
ggsave(filename = "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Plots/volcano_plot_SFTPOM_coreom.svg",
       plot = plot, 
       width = 6, height = 6, dpi = 300, device = "svg")
```

```{r}
# Print OTUIDs of Core Taxa (Circled)
core_otus <- data %>%
  filter(Core_Taxa == "Core") %>%
  select(OTUID)

# View the OTUIDs that were circled
print(core_otus)

# Count how many OTUs were circled
cat("Number of Core Taxa (Circled):", nrow(core_otus), "\n")

# Save the OTUIDs to a CSV file (optional)
write.csv(core_otus, "core_OTUIDs_sftpOM.csv", row.names = FALSE)

print(non_highlighted_otus)

# Save to CSV (optional)
write.csv(non_highlighted_otus, "non_highlighted_OTUIDs_sftpOM.csv", row.names = FALSE)
```








**Making volcano plot for SF vs SJ OM**


```{r}
# Read in main dataset
data <- read_excel("/Users/kyasparks/Desktop/MaaslinResults_includingNonSignif_16S_corrected.xlsx", sheet = "SJ v SF OM") %>% as.data.frame()

# Read SF5 discriminant OTUs
discriminant_otus_SFOM <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrim_results2.xlsx", sheet = "SFOM_discrim") %>%
  select(OTUID) %>%
  pull()

# Read TP5 discriminant OTUs
discriminant_otus_SJOM <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrim_results2.xlsx", sheet = "SJOM_discrim") %>%
  select(OTUID) %>%
  pull()

# Combine both discriminant OTU lists
discriminant_otus <- unique(c(discriminant_otus_SFOM, discriminant_otus_SJOM))

# Read in core taxa from d70 (core across all sites and depths)
core_taxa_data <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_CoreTaxaAnalysis/CoreTaxa_70_16S_ASCC.xlsx", sheet = "d70_OM") %>%
  select(OTUID)  # Keep taxonomy info  # Keep taxonomy info
```


```{r}
# Extract OTUIDs of core taxa
core_taxa <- core_taxa_data$OTUID

# Ensure matching format
data$OTUID <- trimws(as.character(data$OTUID))
discriminant_otus <- trimws(as.character(discriminant_otus))
core_taxa <- trimws(as.character(core_taxa))
```

```{r}
# Convert discriminant OTUs into a dataframe
discriminant_otus_df <- data.frame(OTUID = discriminant_otus)

# Merge with data correctly
data <- data %>%
  left_join(discriminant_otus_df, by = "OTUID") %>%  # Join properly formatted discriminant OTUs
  left_join(core_taxa_data, by = "OTUID") %>%  # Add taxonomy info
  mutate(
    log_q = -log10(qval),  # Standard -log10(qval) transformation
    Enrichment = case_when(
      qval >= 0.05 ~ "Non-Significant",
      coef > 0 ~ "SF Enriched",
      coef < 0 ~ "SJ Enriched"
    ),
    
    # ✅ Only highlight if the OTU is discriminant AND significant (qval < 0.05)
    Highlight = ifelse(qval < 0.05 & OTUID %in% discriminant_otus, 
                       "Discriminant Across All Sites in OM", "Regular"),
    
    # Core taxa logic
    Core_Taxa = ifelse(OTUID %in% core_taxa, "Core", NA)  
  ) %>%
  mutate(
    # Ensure factors are correctly ordered
    Highlight = factor(Highlight, levels = c("Discriminant Across All Sites in OM", "Regular")),
    Core_Taxa = factor(Core_Taxa, levels = c("Core", NA))
  )
```

```{r}
# Identify OTUs that are significant but NOT in discriminant_otus
non_highlighted_otus <- data %>%
  filter(qval < 0.05 & !(OTUID %in% discriminant_otus))

# Base volcano plot
plot <- ggplot(data, aes(x = coef, y = log_q)) +
  
  # Main points with original enrichment color
  geom_point(aes(color = Enrichment, alpha = Enrichment), size = 3) +
  
  # Custom color scheme
  scale_color_manual(values = c(
    "SF Enriched" = "#A1A481",
    "SJ Enriched" = "#BD7C96",
    "Non-Significant" = "#EEC7A0"
  )) +

  # Adjust alpha for non-significant points
  scale_alpha_manual(values = c("Non-Significant" = 1, "SF Enriched" = 1, "SJ Enriched" = 1)) +
  
  # Overlay discriminant taxa (maintain enrichment color, add black border)
  geom_point(data = data %>% filter(Highlight == "Discriminant Across All Sites in OM"),
             aes(x = coef, y = log_q, fill = Enrichment),  
             shape = 21, size = 3, stroke = 0.4, color = "black") +  

  # Overlay **significant but non-discriminant OTUs** (e.g., red outline)
  geom_point(data = non_highlighted_otus, 
             aes(x = coef, y = log_q, fill = Enrichment),
             shape = 21, size = 3, stroke = 0.6, color = "red") +  

  # Overlay core taxa (circled points)
  geom_point(data = data %>% filter(!is.na(Core_Taxa)),
             aes(x = coef, y = log_q), shape = 21, fill = NA, size = 3, stroke = 1, color = "green") +

  geom_text_repel(data = data %>% filter(!is.na(Core_Taxa)), 
                  aes(x = coef, y = log_q, label = OTUID), 
                  size = 3, box.padding = 0.5, segment.color = "gray") +
  
  
  # Apply same colors to `fill`
  scale_fill_manual(values = c(
    "SF Enriched" = "#A1A481",
    "SJ Enriched" = "#BD7C96",
    "Non-Significant" = "#EEC7A0"
  )) +
  
  # Threshold lines
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", color = "gray") +

  # Labels & theme
  labs(x = "Coefficient", 
       y = "-log10(q-value)", 
       color = "Enrichment", 
       fill = "Enrichment", 
       shape = "Highlighted Taxa") +  

  # Ensure shape legend only applies to discriminant taxa
  scale_shape_manual(values = c("Discriminant Across All Sites in OM" = 21)) +  

  theme_minimal() +
  theme(legend.position = "top")

# Show plot
print(plot)
```


```{r}
# Save the volcano plot as a square SVG file
ggsave(filename = "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Plots/volcano_plot_SFSJOM_coreom.svg",
       plot = plot, 
       width = 6, height = 6, dpi = 300, device = "svg")
```

```{r}
# Print OTUIDs of Core Taxa (Circled)
core_otus <- data %>%
  filter(Core_Taxa == "Core") %>%
  select(OTUID)

# View the OTUIDs that were circled
print(core_otus)

# Count how many OTUs were circled
cat("Number of Core Taxa (Circled):", nrow(core_otus), "\n")

# Save the OTUIDs to a CSV file (optional)
write.csv(core_otus, "core_OTUIDs_SFSJOM.csv", row.names = FALSE)

print(non_highlighted_otus)

# Save to CSV (optional)
write.csv(non_highlighted_otus, "non_highlighted_OTUIDs_sfsjOM.csv", row.names = FALSE)
```


**Making volcano plot for TP vs SJ OM**


```{r}
# Read in main dataset
data <- read_excel("/Users/kyasparks/Desktop/MaaslinResults_includingNonSignif_16S_corrected.xlsx", sheet = "TP v SJ OM") %>% as.data.frame()

# Read SF5 discriminant OTUs
discriminant_otus_TPOM <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrim_results2.xlsx", sheet = "TPOM_discrim") %>%
  select(OTUID) %>%
  pull()

# Read TP5 discriminant OTUs
discriminant_otus_SJOM <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrim_results2.xlsx", sheet = "SJOM_discrim") %>%
  select(OTUID) %>%
  pull()

# Combine both discriminant OTU lists
discriminant_otus <- unique(c(discriminant_otus_TPOM, discriminant_otus_SJOM))

# Read in core taxa from d70 (core across all sites and depths)
core_taxa_data <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_CoreTaxaAnalysis/CoreTaxa_70_16S_ASCC.xlsx", sheet = "d70_OM") %>%
  select(OTUID)  # Keep taxonomy info  # Keep taxonomy info
```


```{r}
# Extract OTUIDs of core taxa
core_taxa <- core_taxa_data$OTUID

# Ensure matching format
data$OTUID <- trimws(as.character(data$OTUID))
discriminant_otus <- trimws(as.character(discriminant_otus))
core_taxa <- trimws(as.character(core_taxa))
```


```{r}
# Convert discriminant OTUs into a dataframe
discriminant_otus_df <- data.frame(OTUID = discriminant_otus)

# Merge with data correctly
data <- data %>%
  left_join(discriminant_otus_df, by = "OTUID") %>%  # Join properly formatted discriminant OTUs
  left_join(core_taxa_data, by = "OTUID") %>%  # Add taxonomy info
  mutate(
    log_q = -log10(qval),  # Standard -log10(qval) transformation
    Enrichment = case_when(
      qval >= 0.05 ~ "Non-Significant",
      coef > 0 ~ "TP Enriched",
      coef < 0 ~ "SJ Enriched"
    ),
    
    # ✅ Only highlight if the OTU is discriminant AND significant (qval < 0.05)
    Highlight = ifelse(qval < 0.05 & OTUID %in% discriminant_otus, 
                       "Discriminant Across All Sites in OM", "Regular"),
    
    # Core taxa logic
    Core_Taxa = ifelse(OTUID %in% core_taxa, "Core", NA)  
  ) %>%
  mutate(
    # Ensure factors are correctly ordered
    Highlight = factor(Highlight, levels = c("Discriminant Across All Sites in OM", "Regular")),
    Core_Taxa = factor(Core_Taxa, levels = c("Core", NA))
  )
```

```{r}
# Identify OTUs that are significant but NOT in discriminant_otus
non_highlighted_otus <- data %>%
  filter(qval < 0.05 & !(OTUID %in% discriminant_otus))

# Base volcano plot
plot <- ggplot(data, aes(x = coef, y = log_q)) +
  
  # Main points with original enrichment color
  geom_point(aes(color = Enrichment, alpha = Enrichment), size = 3) +
  
  # Custom color scheme
  scale_color_manual(values = c(
    "TP Enriched" = "#A1A481",
    "SJ Enriched" = "#BD7C96",
    "Non-Significant" = "#EEC7A0"
  )) +

  # Adjust alpha for non-significant points
  scale_alpha_manual(values = c("Non-Significant" = 1, "TP Enriched" = 1, "SJ Enriched" = 1)) +
  
  # 🔥 Overlay discriminant taxa (maintain enrichment color, add black border)
  geom_point(data = data %>% filter(Highlight == "Discriminant Across All Sites in OM"),
             aes(x = coef, y = log_q, fill = Enrichment),  
             shape = 21, size = 3, stroke = 0.4, color = "black") +  

  # Overlay **significant but non-discriminant OTUs** (e.g., red outline)
  geom_point(data = non_highlighted_otus, 
             aes(x = coef, y = log_q, fill = Enrichment),
             shape = 21, size = 3, stroke = 0.6, color = "red") +  

  # Overlay core taxa (circled points)
  geom_point(data = data %>% filter(!is.na(Core_Taxa)),
             aes(x = coef, y = log_q), shape = 21, fill = NA, size = 3, stroke = 1, color = "green") +

  geom_text_repel(data = data %>% filter(!is.na(Core_Taxa)), 
                  aes(x = coef, y = log_q, label = OTUID), 
                  size = 3, box.padding = 0.5, segment.color = "gray") +
  
  
  # Apply same colors to `fill`
  scale_fill_manual(values = c(
    "TP Enriched" = "#A1A481",
    "SJ Enriched" = "#BD7C96",
    "Non-Significant" = "#EEC7A0"
  )) +
  
  # Threshold lines
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  geom_hline(yintercept = -log10(0.05), linetype = "dotted", color = "gray") +

  # Labels & theme
  labs(x = "Coefficient", 
       y = "-log10(q-value)", 
       color = "Enrichment", 
       fill = "Enrichment", 
       shape = "Highlighted Taxa") +  

  # Ensure shape legend only applies to discriminant taxa
  scale_shape_manual(values = c("Discriminant Across All Sites in OM" = 21)) +  

  theme_minimal() +
  theme(legend.position = "top")

# Show plot
print(plot)
```

```{r}
# Save the volcano plot as a square SVG file
ggsave(filename = "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Plots/volcano_plot_TPSJOM.svg",
       plot = plot, 
       width = 6, height = 6, dpi = 300, device = "svg")
```

```{r}
# Print OTUIDs of Core Taxa (Circled)
core_otus <- data %>%
  filter(Core_Taxa == "Core") %>%
  select(OTUID)

# View the OTUIDs that were circled
print(core_otus)

# Count how many OTUs were circled
cat("Number of Core Taxa (Circled):", nrow(core_otus), "\n")

# Save the OTUIDs to a CSV file (optional)
write.csv(core_otus, "core_OTUIDs_tpsjOM.csv", row.names = FALSE)

print(non_highlighted_otus)

# Save to CSV (optional)
write.csv(non_highlighted_otus, "non_highlighted_OTUIDs_tpsjOM_coreom.csv", row.names = FALSE)

highlighted_otus <- data %>% 
  filter(Highlight == "Discriminant Across All Sites in OM") %>% 
  select(OTUID)

print(highlighted_otus)

write.csv(highlighted_otus, "highlighted_otus_tpsjOM.csv", row.names = FALSE)
```