```{r}
library(ggplot2)
library(tidyverse)
library(phyloseq)
library(dplyr)
library(readxl)
library(ggrepel)
library(openxlsx)
```

# Reading in the Maaslin2 results
## Shallow 0-5 cm only

```{r}
data_SJSF_shallow <- read_excel("/Users/kyasparks/Desktop/MaaslinResults_includingNonSignif_ITS_corrected.xlsx", sheet = "SF v SJ shallow")

data_SFTP_shallow <- read_excel("/Users/kyasparks/Desktop/MaaslinResults_includingNonSignif_ITS_corrected.xlsx", sheet = "SF v TP shallow")

data_SJTP_shallow <- read_excel("/Users/kyasparks/Desktop/MaaslinResults_includingNonSignif_ITS_corrected.xlsx", sheet = "TP v SJ shallow")
```

## Deep 5-15 cm only

```{r}
data_SJSF_deep <- read_excel("/Users/kyasparks/Desktop/MaaslinResults_includingNonSignif_ITS_corrected.xlsx", sheet = "SJ v SF deep")

data_SFTP_deep <- read_excel("/Users/kyasparks/Desktop/MaaslinResults_includingNonSignif_ITS_corrected.xlsx", sheet = "SF v TP deep")

data_SJTP_deep <- read_excel("/Users/kyasparks/Desktop/MaaslinResults_includingNonSignif_ITS_corrected.xlsx", sheet = "TP v SJ deep")
```


## Organic matter only

```{r}
data_SJSF_OM <- read_excel("/Users/kyasparks/Desktop/MaaslinResults_includingNonSignif_ITS_corrected.xlsx", sheet = "SJ v SF OM")

data_SFTP_OM<- read_excel("/Users/kyasparks/Desktop/MaaslinResults_includingNonSignif_ITS_corrected.xlsx", sheet = "SF v TP OM")

data_SJTP_OM <- read_excel("/Users/kyasparks/Desktop/MaaslinResults_includingNonSignif_ITS_corrected.xlsx", sheet = "TP v SJ OM")
```



# Inner joining Maaslin comparisons to figure out which taxa are discrimninant for each site
## For example: for the  0-5 cm I want to know which taxa are discriminant and conserved when looking at Taylor Park, so I need to look at the OTUIDs that are consistent across all comparisons that include Taylor Park (TP vs SF and TP vs SJ)


```{r}
data_SF5_discrim<-inner_join(data_SFTP_shallow,data_SJSF_shallow,by="feature")
data_TP5_discrim<-inner_join(data_SFTP_shallow,data_SJTP_shallow,by="feature")
data_SJ5_discrim<-inner_join(data_SJSF_shallow,data_SJTP_shallow,by="feature")
```

```{r}
data_SF15_discrim<-inner_join(data_SFTP_deep,data_SJSF_deep,by="feature")
data_TP15_discrim<-inner_join(data_SFTP_deep,data_SJTP_deep,by="feature")
data_SJ15_discrim<-inner_join(data_SJSF_deep,data_SJTP_deep,by="feature")
```

```{r}
data_SFOM_discrim<-inner_join(data_SFTP_OM,data_SJSF_OM,by="feature")
data_TPOM_discrim<-inner_join(data_SFTP_OM, data_SJTP_OM,by="feature")
data_SJOM_discrim<-inner_join(data_SJSF_OM,data_SJTP_OM,by="feature")
```


```{r}
library(openxlsx)

# Create a new workbook
wb <- createWorkbook()

# Add sheets to the workbook
addWorksheet(wb, "SF5_discrim")
addWorksheet(wb, "TP5_discrim")
addWorksheet(wb, "SJ5_discrim")
addWorksheet(wb, "SF15_discrim")
addWorksheet(wb, "TP15_discrim")
addWorksheet(wb, "SJ15_discrim")
addWorksheet(wb, "SFOM_discrim")
addWorksheet(wb, "TPOM_discrim")
addWorksheet(wb, "SJOM_discrim")

# Write data to the sheets
writeData(wb, "SF5_discrim", data_SF5_discrim)
writeData(wb, "TP5_discrim", data_TP5_discrim)
writeData(wb, "SJ5_discrim", data_SJ5_discrim)
writeData(wb, "SF15_discrim", data_SF15_discrim)
writeData(wb, "TP15_discrim", data_TP15_discrim)
writeData(wb, "SJ15_discrim", data_SJ15_discrim)
writeData(wb, "SFOM_discrim", data_SFOM_discrim)
writeData(wb, "TPOM_discrim", data_TPOM_discrim)
writeData(wb, "SJOM_discrim", data_SJOM_discrim)

# Save the workbook
saveWorkbook(wb, "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/ITS/ITS_Maaslin/Maaslin_discrim_results_ITS.xlsx", overwrite = TRUE)
```


```{r}
data_SFTP_shallow <- data_SFTP_shallow %>%
  group_by(OTUID) %>%
  summarize(
    coef_SFTP = median(coef, na.rm = TRUE),
    qval_SFTP = median(qval, na.rm = TRUE),
    pval_SFTP = median(pval, na.rm = TRUE),
    .groups = "drop"
  )

data_SJSF_shallow <- data_SJSF_shallow %>%
  group_by(OTUID) %>%
  summarize(
    coef_SJSF = median(coef, na.rm = TRUE),
    qval_SJSF = median(qval, na.rm = TRUE),
    pval_SJSF = median(pval, na.rm = TRUE),
    .groups = "drop"
  )

data_SF5_discrim <- inner_join(data_SFTP_shallow, data_SJSF_shallow, by = "OTUID")
```

```{r}
coef_SFTP = median(coef, na.rm = TRUE)
```

```{r}
colnames(data_SFTP_shallow)
colnames(data_SJSF_shallow)
```

# Inner joining Maaslin comparisons to figure out which taxa are discrimninant for each site
## For example: for the  0-5 cm I want to know which taxa are discriminant and conserved when looking at Taylor Park, so I need to look at the OTUIDs that are consistent across all comparisons that include Taylor Park (TP vs SF and TP vs SJ)


**CORRECTED**
### Actually considering coefficients
#### sf v. sj shallow: when the coef is negative it is discriminant for SF and when its positive its discriminant for SJ
#### sf v. tp shallow: when the coef is negative it is discriminant for SF and when its positive its discriminant for TP
#### tp v. sj shallow: when the coef is negative it is discriminant for SJ and when its positive its discriminant for TP
#### sj v. sf deep: when the coef is negative it is discriminant for SF and when its positive its discriminant for SJ
#### sf v. tp deep: when the coef is negative it is discriminant for SF and when its positive its discriminant for TP
#### tp v. sj deep: when the coef is negative it is discriminant for SJ and when its positive its discriminant for TP
#### sj v. sf OM: when the coef is negative it is discriminant for SF and when its positive its discriminant for SJ
#### sf v. tp OM: when the coef is negative it is discriminant for SF and when its positive its discriminant for TP
#### tp v. sj OM: when the coef is negative it is discriminant for SJ and when its positive its discriminant for TP


```{r}
library(readxl)
library(dplyr)
library(openxlsx)

# Load data
data_SJSF_shallow <- read_excel("/Users/kyasparks/Desktop/MaaslinResults_includingNonSignif_ITS_corrected.xlsx", sheet = "SJ v SF shallow")
data_SFTP_shallow <- read_excel("/Users/kyasparks/Desktop/MaaslinResults_includingNonSignif_ITS_corrected.xlsx", sheet = "SF v TP shallow")
data_SJTP_shallow <- read_excel("/Users/kyasparks/Desktop/MaaslinResults_includingNonSignif_ITS_corrected.xlsx", sheet = "TP v SJ shallow")

data_SJSF_deep <- read_excel("/Users/kyasparks/Desktop/MaaslinResults_includingNonSignif_ITS_corrected.xlsx", sheet = "SJ v SF deep")
data_SFTP_deep <- read_excel("/Users/kyasparks/Desktop/MaaslinResults_includingNonSignif_ITS_corrected.xlsx", sheet = "SF v TP deep")
data_SJTP_deep <- read_excel("/Users/kyasparks/Desktop/MaaslinResults_includingNonSignif_ITS_corrected.xlsx", sheet = "TP v SJ deep")

data_SJSF_OM <- read_excel("/Users/kyasparks/Desktop/MaaslinResults_includingNonSignif_ITS_corrected.xlsx", sheet = "SJ v SF OM")
data_SFTP_OM <- read_excel("/Users/kyasparks/Desktop/MaaslinResults_includingNonSignif_ITS_corrected.xlsx", sheet = "SF v TP OM")
data_SJTP_OM <- read_excel("/Users/kyasparks/Desktop/MaaslinResults_includingNonSignif_ITS_corrected.xlsx", sheet = "TP v SJ OM")

# Function to classify discriminant OTUs
classify_discriminant <- function(data, group1, group2) {
  data %>%
    filter(qval < 0.05) %>%  # Consider only significant OTUs
    mutate(discriminant_for = ifelse(coef < 0, group1, group2))
}

# Apply classification
data_SJSF_shallow <- classify_discriminant(data_SJSF_shallow, "SF", "SJ")
data_SFTP_shallow <- classify_discriminant(data_SFTP_shallow, "SF", "TP")
data_SJTP_shallow <- classify_discriminant(data_SJTP_shallow, "SJ", "TP")

data_SJSF_deep <- classify_discriminant(data_SJSF_deep, "SF", "SJ")
data_SFTP_deep <- classify_discriminant(data_SFTP_deep, "SF", "TP")
data_SJTP_deep <- classify_discriminant(data_SJTP_deep, "SJ", "TP")

data_SJSF_OM <- classify_discriminant(data_SJSF_OM, "SF", "SJ")
data_SFTP_OM <- classify_discriminant(data_SFTP_OM, "SF", "TP")
data_SJTP_OM <- classify_discriminant(data_SJTP_OM, "SJ", "TP")

# Inner join to find consistent discriminant OTUs
data_SF5_discrim <- inner_join(data_SFTP_shallow, data_SJSF_shallow, by = "feature")
data_TP5_discrim <- inner_join(data_SFTP_shallow, data_SJTP_shallow, by = "feature")
data_SJ5_discrim <- inner_join(data_SJSF_shallow, data_SJTP_shallow, by = "feature")

data_SF15_discrim <- inner_join(data_SFTP_deep, data_SJSF_deep, by = "feature")
data_TP15_discrim <- inner_join(data_SFTP_deep, data_SJTP_deep, by = "feature")
data_SJ15_discrim <- inner_join(data_SJSF_deep, data_SJTP_deep, by = "feature")

data_SFOM_discrim <- inner_join(data_SFTP_OM, data_SJSF_OM, by = "feature")
data_TPOM_discrim <- inner_join(data_SFTP_OM, data_SJTP_OM, by = "feature")
data_SJOM_discrim <- inner_join(data_SJSF_OM, data_SJTP_OM, by = "feature")

# Save results to Excel
wb <- createWorkbook()

# Add sheets
addWorksheet(wb, "SF5_discrim")
addWorksheet(wb, "TP5_discrim")
addWorksheet(wb, "SJ5_discrim")
addWorksheet(wb, "SF15_discrim")
addWorksheet(wb, "TP15_discrim")
addWorksheet(wb, "SJ15_discrim")
addWorksheet(wb, "SFOM_discrim")
addWorksheet(wb, "TPOM_discrim")
addWorksheet(wb, "SJOM_discrim")

# Write data
writeData(wb, "SF5_discrim", data_SF5_discrim)
writeData(wb, "TP5_discrim", data_TP5_discrim)
writeData(wb, "SJ5_discrim", data_SJ5_discrim)
writeData(wb, "SF15_discrim", data_SF15_discrim)
writeData(wb, "TP15_discrim", data_TP15_discrim)
writeData(wb, "SJ15_discrim", data_SJ15_discrim)
writeData(wb, "SFOM_discrim", data_SFOM_discrim)
writeData(wb, "TPOM_discrim", data_TPOM_discrim)
writeData(wb, "SJOM_discrim", data_SJOM_discrim)

# Save workbook
saveWorkbook(wb, "/Users/kyasparks/Desktop/Maaslin_discrim_results_corrected_ITS.xlsx", overwrite = TRUE)
```