
# Inner joining Maaslin comparisons to figure out which taxa are discrimninant for each site
## For example: for the  0-5 cm I want to know which taxa are discriminant and conserved when looking at Taylor Park, so I need to look at the OTUIDs that are consistent across all comparisons that include Taylor Park (TP vs SF and TP vs SJ)


```{r}
library(ggplot2)
library(tidyverse)
library(phyloseq)
library(dplyr)
library(readxl)
library(ggrepel)
library(openxlsx)
```

# Reading in the Maaslin2 results
## Shallow 0-5 cm only

```{r}
data_SJSF_shallow <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_allresults_COMBINED_16S.xlsx", sheet = "SJ v SF shallow")
```

```{r}
data_SFTP_shallow <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_allresults_COMBINED_16S.xlsx", sheet = "SF v TP shallow")
```


```{r}
data_SJTP_shallow <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_allresults_COMBINED_16S.xlsx", sheet = "TP v SJ shallow")
```



## Deep 5-15 cm only

```{r}
data_SJSF_deep <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_allresults_COMBINED_16S.xlsx", sheet = "SJ v SF deep")
```

```{r}
data_SFTP_deep <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_allresults_COMBINED_16S.xlsx", sheet = "SF v TP deep")
```

```{r}
data_SJTP_deep <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_allresults_COMBINED_16S.xlsx", sheet = "TP v SJ deep")
```



## Organic matter only

```{r}
data_SJSF_OM <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_allresults_COMBINED_16S.xlsx", sheet = "SJ v SF OM")
```

```{r}
data_SFTP_OM<- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_allresults_COMBINED_16S.xlsx", sheet = "SF v TP OM")
```



```{r}
data_SJTP_OM <- read_excel("/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_allresults_COMBINED_16S.xlsx", sheet = "TP v SJ OM")
```


# Inner joining Maaslin comparisons to figure out which taxa are discrimninant for each site
## For example: for the  0-5 cm I want to know which taxa are discriminant and conserved when looking at Taylor Park, so I need to look at the OTUIDs that are consistent across all comparisons that include Taylor Park (TP vs SF and TP vs SJ)

```{r}
data_SF5_discrim <- inner_join(data_SFTP_shallow, data_SJSF_shallow, by = "OTUID")
data_TP5_discrim <- inner_join(data_SFTP_shallow, data_SJTP_shallow, by = "OTUID")
data_SJ5_discrim <- inner_join(data_SJSF_deep, data_SJTP_deep, by = "OTUID")
```

```{r}
data_SF15_discrim <- inner_join(data_SFTP_deep,data_SJSF_deep, by =c("OTUID"="OTUID"))
```

```{r}
data_TP15_discrim <- inner_join(data_SFTP_deep,data_SJTP_deep, by =c("OTUID"="OTUID"))
```

```{r}
data_SJ15_discrim <- inner_join(data_SJSF_deep,data_SJTP_deep, by =c("OTUID"="OTUID"))
```

```{r}
data_SFOM_discrim <- inner_join(data_SFTP_shallow,data_SJSF_OM, by =c("OTUID"="OTUID"))
```

```{r}
data_TPOM_discrim <- inner_join(data_SFTP_shallow,data_SJTP_OM, by =c("OTUID"="OTUID"))
```

```{r}
data_SJOM_discrim <- inner_join(data_SJSF_shallow,data_SJTP_OM, by =c("OTUID"="OTUID"))
```

```{r}
library(openxlsx)

# Create a new workbook
wb <- createWorkbook()

# Add sheets to the workbook
addWorksheet(wb, "SF5_discrim")
addWorksheet(wb, "TP5_discrim")
addWorksheet(wb, "SJ5_discrim")
addWorksheet(wb, "SF15_discrim")
addWorksheet(wb, "TP15_discrim")
addWorksheet(wb, "SJ15_discrim")
addWorksheet(wb, "SFOM_discrim")
addWorksheet(wb, "TPOM_discrim")
addWorksheet(wb, "SJOM_discrim")

# Write data to the sheets
writeData(wb, "SF5_discrim", data_SF5_discrim)
writeData(wb, "TP5_discrim", data_TP5_discrim)
writeData(wb, "SJ5_discrim", data_SJ5_discrim)
writeData(wb, "SF15_discrim", data_SF15_discrim)
writeData(wb, "TP15_discrim", data_TP15_discrim)
writeData(wb, "SJ15_discrim", data_SJ15_discrim)
writeData(wb, "SFOM_discrim", data_SFOM_discrim)
writeData(wb, "TPOM_discrim", data_TPOM_discrim)
writeData(wb, "SJOM_discrim", data_SJOM_discrim)

# Save the workbook
saveWorkbook(wb, "/Users/kyasparks/Desktop/CSU 2023-2024/Projects/ASCC/ASCC_July2024_USETHIS_ks/16S/16S_Maaslin/Maaslin_discrim_results.xlsx", overwrite = TRUE)
```